\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Introduction}
\label{chap:intro}
% ------------------------------------------------------------
\section{Side-Channel Attacks and the Masking Countermeasure}
\label{sec:intro-masking}

A cryptographic scheme can be secure on paper and insecure on the bench. When an
algorithm runs on a real device its power draw, its timing, and its
electromagnetic emanations all vary with the data it handles, and an adversary who
measures those physical quantities can recover a key the mathematics was supposed
to protect. This is called a \emph{side-channel attack}, and differential power analysis
\parencite{kocher1999differential} is a canonical example. By correlating many
power traces against a guess of an intermediate value, it extracts secret keys from
implementations that are deemed secure in theory. So, the insecurity
lies not in the cipher but in the way the hardware computes it.

\emph{Masking} is the standard algorithmic countermeasure against such attacks
\parencite{chari1999towards}. It splits each secret value into $n$ \emph{shares} so
that any $n-1$ of them, taken together, are independent of the secret. Then,
computations are carried out on the shares rather than on the secret itself. The
shares are recombined by an operation fixed in advance, and the choice of operation
names the scheme: arithmetic masking adds the shares in a ring and multiplicative
masking multiplies them, but the form we adopt is \emph{Boolean masking}, where the
operation is exclusive-or over $\mathbb{F}_2$. A secret $x$ is written as $x = x_1
\oplus \dots \oplus x_n$ with the first $n-1$ shares drawn uniformly at random, so
that any $n-1$ of them reveal nothing about $x$.

The reason this helps countering side-channel attacks is physical.
Each share leaks separately and noisily, and a
measurement that reflects only some of the shares of a secret is, by construction,
independent of that secret; it carries no information an attacker can use. To learn
anything the attacker must combine the leakage of several shares at once and
average away the noise, and the cost of doing so grows steeply. For a fixed noise
level, the number of traces needed to execute a successful attack grows exponentially
in the number of shares \parencite{chari1999towards}. Adding a share is therefore a
tunable defence as well, trading circuit size for attack cost.

The difficulty is that this gurantee holds only if the shares are kept apart
at places the attacker can reach. Implementing masking across a whole cipher is
delicate, because the non-linear operations of the unmasked algorithm force the
shares of different variables to interact in the masked version, and it is precisely
those interactions that a scheme must handle with care \parencite{10.1007/978-3-540-45146-4_27,
rivain2010provably}. Simple mistakes can reopen the leak, which masking was meant to
close.

\section{From Physical Leakage to a Probing Adversary}
\label{sec:intro-model}

Reasoning about masking directly in terms of noisy power traces is unwieldy. The
adversary's advantage is a statistical quantity that depends on the device, the
measurement setup, and the noise distribution, none of which the designers
control completely. The field instead reasons about masking through a discrete abstraction,
called the \emph{probing model} owing to Ishai, Sahai, and Wagner
\parencite{10.1007/978-3-540-45146-4_27}. In this model an adversary of order $d$
does not measure anything; it simply chooses any $d$ wires of the circuit and is
handed their exact values. The circuit is
\emph{$d$-probing secure} when no such choice of $d$ wires lets the adversary tell
one secret from another. An order-$d$ masking accordingly uses $d+1$ shares, so
that no $d$ probes ever meet all the shares of a secret at once.

The probing model trades physical fidelity for a question one can actually decide.
Security becomes a combinatorial property of the circuit, checkable in principle by
inspecting finitely many sets of wires. What makes the trade legitimate is that the
abstraction is not merely convenient but provably faithful. Duc, Dziembowski, and
Faust \parencite{duc2014unifying} gave a reduction showing that security in the
probing model implies security in the more realistic \emph{noisy leakage} model, in
which every wire leaks a noisy function of its value; the abstract, value-based
guarantee thus entails the physical one. We should note that the reduction is not trivial,
it asks that the noise grow with the number of shares, and it carries a dependence on the field
size, but establishing it as a question about leakage models rather than about any
particular circuit, is useful in practice. We therefore take
$d$-probing security as our target throughout and do not revisit this bridge. The
reader should simply keep in mind that a $d$-probing-secure circuit is what a
faithful masked implementation is supposed to achieve.

\section{Verifying Masked Implementations}
\label{sec:intro-verification}

Masking is easy to get subtly wrong, and the probing model makes the failures
precise. A single wire may recombine several
shares of one secret, or a random mask meant to hide one value may cancel against a
copy of itself elsewhere; either case reopens the very leak the masking was meant to
close. Correctness cannot be judged wire by wire as well, because $d$-probing security
is a statement about how wires leak \emph{jointly}, and the number of $d$-subsets
grows quickly with the size of the circuit and the order $d$, which puts checking by
hand out of reach for all but the smallest implementations.

What one wants, then, is a tool that takes a masked circuit and decides,
mechanically and for all $d$-subsets at once, whether any of them leaks. This is the
problem of \emph{formally verifying} a masked implementation, and a line of work has
grown around it \parencite{barthe2015verified, 10.1007/978-3-030-29959-0_15}. Two
difficulties stand in the way. Even a single subset already poses a question about a
distribution over all of the circuit's internal random choices
and there are combinatorially many subsets to settle. The tools
that scale do not compute those distributions explicitly.

\section{Scope and Threat Model}
\label{sec:intro-scope}

Before moving on, we should fix the model precisely. The circuit definition we work with
is arithmetic circuits over $\mathbb{F}_2$,
built from two-input XOR and AND gates, and its inputs are partitioned into secret,
random, and public wires. Leakage is \emph{value-based}, so, a probed wire leaks the
logical value it carries and nothing more. Physical effects that make a wire leak
more than its final value --- such as glitches and transitions in hardware
\parencite{10.1007/978-3-030-29959-0_15} --- are outside the model, as is any question
of composing separately verified circuits into a larger one. The object of study is
the exact probing security of a single circuit, decided over its wire values.

Within that model our aim is exact verification rather than a sound approximation.
So, when the tool reports a circuit secure, every one of its $\sum_{i \le d}
\binom{|\mathcal{W}|}{i}$ observation sets has been accounted for; when it reports a
circuit insecure it returns a specific leaking observation as a witness --- with the
exception of the incomplete fragment flagged in
Chapter~\ref{chap:design}. % TODO: do we really flag this!
% The decision problem underlying even a single observation
% is hard in general, in a sense Chapter~\ref{chap:couplings} makes precise, and it is
% that hardness that forces the constructive route the tool takes rather than a direct
% count.

\section{Contributions}
\label{sec:intro-contributions} 

This thesis presents \toolname\footnote{\url{https://github.com/Sabanci-Theory/coppula/}}, a verifier of $d$-probing security for masked
arithmetic circuits over $\mathbb{F}_2$, implemented in Lean~4. Its design rests on
three ideas. Before developing them we place the verification problem itself in the
complexity hierarchy, and the contributions follow in that order.

\textbf{Placing the exact-verification problem.} Before building anything we ask
how hard the problem really is. The law of a probe is actually a histogram of integer tape counts,
and we show that computing one such count is $\#\mathrm{P}$-complete, while deciding
whether two secrets leave the law unchanged compares two such counts and is
$\mathrm{C}_{=}\mathrm{P}$-complete (Chapter~\ref{chap:couplings}). Therefore, even a single
probe lands on a level of the complexity hierarchy above the polynomial hierarchy,
which is what rules out exact counting and motivates the constructive
route the rest of our tool takes.

\textbf{Deciding one observation by construction.} At the heart of the tool is a
rewrite engine that decides whether a single probe depends on the secret by
substituting its randoms away one at a time, driving the observation toward a
secret-free form (Chapter~\ref{chap:design}). The engine is layered as a cheap
single-occurrence rule first, followed by a complete substitution loop behind it,
and we prove its verdicts sound. Some masks, though, are buried inside products where ordinary
substitution cannot reach them, and to expose those we cleverly add \emph{multiplicative
factoring} steps that reshape products so the hidden randoms surface.

\textbf{Covering every observation without enumerating them.} Certifying each
$d$-subset in turn is infeasible, so the tool instead splits the space of
observations recursively, in the manner of \texttt{maskVerif}
\parencite{10.1007/978-3-030-29959-0_15}, and tries to discharge whole regions at once.
The delicate point, of course, is that the split must lose nothing. So, we prove it covers every
observation exactly once, by reading a Vandermonde-style subset-counting identity at
the level of sets (Chapter~\ref{chap:design}).

\textbf{Reading a proof as a coupling, and growing the probe.} A successful
discharge can be read as more than a yes-or-no verdict. We reinterpret it as a
measure-preserving \emph{coupling} of the circuit's randomness, in the tradition of
coupling proofs for probabilistic programs \parencite{barthe2017coupling}, and
develop this reading for probing security in Chapter~\ref{chap:couplings}. The
benefit is that one certificate can apply for many wires, in turn we grow a certified probe
into a whole \emph{safe set} of wires that a single discharge settles. We study three
mechanisms for this growth, and adopt the coupling-based one, proving that the other
two sit inside it (Chapter~\ref{chap:design}).

\textbf{Implementation and evaluation.} We describe the data structures that make
the engine efficient (a hash-consed circuit representation and a bit-sliced test for the extension)
(Chapter~\ref{chap:implementation}). and we evaluate
\toolname\ on a suite of masked gadgets, reporting its verification outcomes
(Chapter~\ref{chap:evaluation}).

\section{Thesis Outline}
\label{sec:intro-outline}

Chapter~\ref{chap:preliminaries} fixes notation and defines arithmetic circuits, the
probing model, and $d$-probing security, and studies the randomness-substitution
technique we rely on throughout. Chapter~\ref{chap:related} surveys the verification
literature we build on and depart from. Chapter~\ref{chap:couplings} develops the
coupling view of probing security and situates the underlying decision problem in
the complexity hierarchy. Chapter~\ref{chap:design}, the core of the thesis,
presents the algorithm with its three components together with their correctness
proofs, and Chapter~\ref{chap:implementation} gives the concrete data structures
behind them. Chapter~\ref{chap:evaluation} reports the experiments.
Chapter~\ref{chap:futurework} outlines the directions our approach does not treat,
and Chapter~\ref{chap:conclusion} concludes the work.
