\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Preliminaries}
\label{chap:preliminaries}
% ------------------------------------------------------------
\section{Notation and Mathematical Background}
\label{sec:notation}

This chapter collects the background the rest of the thesis builds on and fixes
the notation for it. We use $\{0,1\}$ or $\mathbb{F}_2$ for the set of
bits. Tuples are set in boldface, $\mathbf{x} = (x_1, \dots, x_n)$, with $|\mathbf{x}|$
denoting their length. For a finite set $X$ we write $\mathcal{P}(X)$ for its power set and
$\binom{X}{k}$ for the family of its $k$-element subsets. Functions are total
unless we say otherwise, and we identify a function $\{0,1\}^n \to \{0,1\}$ with the
Boolean function it computes. Notation that is used only locally is introduced
where it appears.
  \oldsubsection{\texorpdfstring%
    {The Field $\mathbb{F}_2$ and Multilinear Boolean Polynomials}%
    {The Field F2 and Multilinear Boolean Polynomials}}
  \label{sec:field-f2}

  The field $\mathbb{F}_2$ has carrier $\{0,1\}$ with addition $\oplus$, the
  exclusive-or, and multiplication $\odot$, the logical-and. Addition is its own
  inverse, so $x \oplus x = 0$, and multiplication is idempotent, $x \odot x = x$.

  Every function $f : \mathbb{F}_2^{\,n} \to \mathbb{F}_2$ is a polynomial over
  $\mathbb{F}_2$, and idempotence lets us take that polynomial to have no variable
  raised above the first power. So $f$ has a unique \emph{multilinear}
  representation,
  \[
    f(x_1, \dots, x_n) \;=\; \bigoplus_{S \subseteq [n]} c_S \bigodot_{i \in S} x_i,
    \qquad c_S \in \mathbb{F}_2,
  \]
  its \emph{algebraic normal form} (ANF). Uniqueness is what turns the ANF into a
  decision procedure for the questions we put to it. In particular, note that, $f$ depends on a
  variable $x_i$ (meaning that flipping $x_i$ can change the output) if and only if
  some monomial of the ANF contains $x_i$.

  \subsection{Probability Distributions and Statistical Independence}
  \label{sec:prob-independence}

  Probing security is a statement about distributions, so we fix the little we
  need. A \emph{distribution} on a finite set $X$ is a function $\mu : X \to [0,1]$
  with $\sum_{x \in X} \mu(x) = 1$, and the \emph{uniform} distribution assigns
  $\mu(x) = 1/|X|$ to every $x$. A random variable $Z$ into a finite set induces
  the distribution $x \mapsto \Pr[Z = x]$, its \emph{law}, and two random variables
  are \emph{identically distributed} when their laws agree on every value.

  The security notion we are after is the independence of a family of laws from a
  parameter. Suppose a random variable $Z$ depends on a parameter $s$, so that each
  value of $s$ gives a law $\mu_s$. We say $Z$ is \emph{independent of} $s$ when
  $\mu_s$ is one and the same distribution for every value of $s$. For probing
  security the parameter is the secret assignment and the law is that of a probe,
  as Section~\ref{sec:d-probing} makes precise.
  One way to prove that two laws are equal is through a measure-preserving
  bijection: if $\pi$ carries the uniform distribution on $X$ to itself, then for
  any function $g$ the variables $g(Z)$ and $g(\pi(Z))$ are identically distributed
  when $Z$ is uniform.

  \subsection{Boolean Circuits and Gates}
  \label{sec:bool-circuits}

We now introduce the object the rest of the thesis is built on.
The definition is the standard digraph formulation of a circuit, specialised to
arithmetic over the two-element field.

\begin{definition}[Arithmetic circuit over $\mathbb{F}_2$]
\label{def:circuit}
Fix the field $\mathbb{F}_2 = \{0,1\}$ and a set $S = \{s_i\}$ of \emph{gate
functions}, each of some arity $a_i \in \mathbb{N}_{\geq 1}$, i.e.\ each $s_i$
maps $\mathbb{F}_2^{\,a_i}$ to $\mathbb{F}_2$. A \emph{circuit} $C$ is a finite
directed graph (a \emph{digraph}) containing no directed cycle, such that:
\begin{itemize}
  \item Every vertex of indegree $0$ is labelled either by an
  \emph{indeterminate}, in which case it is an \emph{input gate}, or by an
  element of $\mathbb{F}_2$, in which case it is a \emph{constant gate}. The
  input gates are partitioned into \emph{secret} inputs $\mathcal{S}$,
  \emph{random} inputs $\mathcal{R}$, and \emph{public} inputs $\mathcal{P}$.
  \item Every remaining vertex is a \emph{computation gate}, labelled by a
  function $s_i \in S$ whose arity $a_i$ coincides with the vertex's indegree.
  A computation gate of outdegree $0$ is an \emph{output gate}.
\end{itemize}
We call the vertices \emph{wires}, and for a computation gate we speak
interchangeably of the gate and the wire carrying its output.
\end{definition}

Over $\mathbb{F}_2$ there is no difference between the Boolean and the arithmetic
view of such a circuit: exclusive-or is addition, logical-and is
multiplication, and the two together already express every function
$\mathbb{F}_2^{\,n} \to \mathbb{F}_2$, i.e. they are universal. So we fix
\[
  S = \{\,\oplus,\ \odot\,\},
\]
addition and multiplication in $\mathbb{F}_2$ (Section~\ref{sec:field-f2}), both
of arity $2$. Every computation gate then has indegree exactly $2$ and reads
\[
  w_i = w_j \oplus w_k \qquad \text{or} \qquad w_i = w_j \odot w_k .
\]
Throughout the thesis $\odot$ denotes multiplication in $\mathbb{F}_2$; we
might sometimes use $\cdot$ for the same operation.

Acyclicity is what makes a circuit computable. It lets us list the wires in a
topological order $w_1, \dots, w_m$ in which the two operands of every gate
appear before the gate itself, so that $j, k < i$. Fixing a value in
$\mathbb{F}_2$ for every input gate --- a constant gate carries its own label
--- then determines the value of every wire by induction along this order. When
the inputs are clear from context we write simply $w_i$ for this value.

\begin{notation}
For a wire $w_i$, $\mathrm{vars}(w_i) \subseteq \mathcal{R}$ is the set of
random inputs its defining sub-circuit depends on: $\mathrm{vars}(w_i) = \{w_i\}$
if $w_i \in \mathcal{R}$, $\mathrm{vars}(w_i) = \emptyset$ if
$w_i \in \mathcal{S} \cup \mathcal{P}$ or $w_i$ is a constant gate, and
$\mathrm{vars}(w_i) = \mathrm{vars}(w_j) \cup \mathrm{vars}(w_k)$ for a gate
with operands $w_j, w_k$.
\end{notation}

% Two conventions, fixed here once, are used throughout. First, a topological
% order realises the circuit as a \emph{straight-line program}: each wire is
% defined at most once, in terms only of wires already defined, so there is no
% branching and no recomputation of a gate under a different name. We pass freely
% between the digraph of Definition~\ref{def:circuit} and this sequential reading,
% whichever is clearer at the moment. Second, the restriction to $2$-ary gates is
% a modelling choice, not a loss of generality: an $n$-ary sum or product is
% always expressible as a chain of $2$-ary gates.

\section{Secret Sharing and Masking}
\label{sec:secret-sharing}

Physical implementations of cryptographic hardware can leak information. Power draw, electromagnetic
emanation, and timing all correlate with the values a device computes on, and an
attacker who measures them can recover a key that is out of cryptographic reach
\parencite{kocher1999differential}. \emph{Masking} is the standard algorithmic
countermeasure \parencite{chari1999towards}. Instead of computing on a secret
directly, one splits it into several \emph{shares} whose joint distribution hides
it, then rewrites the computation to run on the shares throughout and reconstruct
only at the very end. The promise is that an attacker who sees few enough
intermediate values learns nothing, because too few shares carry no information
about the secret.

  \subsection{Boolean Masking}
  \label{sec:boolean-masking}

  We use \emph{Boolean} masking, where shares combine by exclusive-or. An
  \emph{$n$-sharing} of a bit $x \in \mathbb{F}_2$ is a tuple
  $(x_0, \dots, x_{n-1})$ with $x_0 \oplus \cdots \oplus x_{n-1} = x$. To draw one,
  sample $x_1, \dots, x_{n-1}$ uniformly and independently and set
  $x_0 = x \oplus x_1 \oplus \cdots \oplus x_{n-1}$. The construction is built so
  that any $n-1$ of the shares are then uniform and independent, hence reveal
  nothing about $x$; only the full set of $n$ shares determines it. This is the
  same statement, for one bit, that Section~\ref{sec:prob-independence} calls
  independence of the secret.

  A masked circuit carries every value in shared form and replaces each operation
  by a \emph{gadget}, a sub-circuit computing a sharing of the result from sharings
  of its inputs. Linear operations are trivial, since $\oplus$ acts share by share.
  Multiplication is where the difficulty and the fresh randomness live, such a gadget
  must combine shares of both inputs without ever forming a value that
  depends on a secret. The masked \textsc{and} of Ishai, Sahai and Wagner
  \parencite{10.1007/978-3-540-45146-4_27} is the archetype, and the circuits we
  verify are built from gadgets of this kind.

\section{The ISW Probing Model}
\label{sec:isw-model}

For the security notion to be precise we need to state an adversary. Ishai, Sahai and Wagner
\parencite{10.1007/978-3-540-45146-4_27} gave the model the thesis works in: the
probing model. An adversary places a bounded number of probes on the wires
of a circuit and reads the values they carry, and the circuit is secure when no
choice of probes tells the adversary anything about the secrets. The model
abstracts a physical side-channel attacker into a clean combinatorial game, and it
is the setting in which masking admits proofs.

The connections between the physical side-channel attack and the probing model, however,
are not exposited in this thesis.


  \subsection{Threshold Probing and Observation Sets}
  \label{sec:t-threshold}

  Fix a circuit with secret inputs $\mathcal{S}$, random inputs $\mathcal{R}$, and
  public inputs $\mathcal{P}$ as in Section~\ref{sec:bool-circuits}. A
  \emph{threshold} adversary of order $d$ chooses an \emph{observation set} of at
  most $d$ wires and learns their joint values. The public inputs are known and the secrets
  are held fixed, so the only randomness in play is the internal randomness
  $\mathcal{R}$, sampled uniformly. An observation set therefore carries a joint
  distribution over the draw of $\rho \in \{0,1\}^{\mathcal{R}}$. We call such a set
  a \emph{probe}, and we read each of its wires as a function of $\rho$.

  \subsection{\NoCaseChange{$d$}-Probing Security}
  \label{sec:d-probing}

  \begin{definition}[$d$-probing security]
  \label{def:d-probing}
  A circuit is \emph{$d$-probing secure} if for every observation set of at most
  $d$ wires the joint distribution of that set, taken over uniform
  $\rho \in \{0,1\}^{\mathcal{R}}$, is independent of the secret inputs in the sense
  of Section~\ref{sec:prob-independence}. Equivalently, the law of the set is one
  and the same for every assignment to $\mathcal{S}$.
  \end{definition}

  The order $d$ counts the probes the adversary is granted, and an $n$-share
  masking is usually meant to withstand $d = n-1$ of them. The definition
  quantifies over sets of wires read from a single run of the circuit, rather than single indivdual wires,
  since a masking leaks only jointly. And it is a worst-case statement over all
  $\sum_{i \le d} \binom{|\mathcal{W}|}{i}$ observation sets of the circuit's wires
  $\mathcal{W}$. Verifying this definition for the circuits we defined is the task
  of the whole thesis.

  \subsection{Probabilistic Non-Interference}
  \label{sec:pni}

  The distributional condition of Definition~\ref{def:d-probing} has an equivalent
  phrasing the verification literature prefers, \emph{probabilistic
  non-interference} \parencite{barthe2015verified}. It helps to study it and understand 
  in different views, so below we give three perspectives that realize the same notion. Fix a
  probe $O$ with $|O| \le d$. Under a secret assignment $\mathbf{a} \in
  \{0,1\}^{\mathcal{S}}$ and uniform randomness $\rho$, the probe takes the joint
  value $O(\rho; \mathbf{a}) = \big(w(\rho; \mathbf{a})\big)_{w \in O}$ and we
  write $\mathcal{L}_{\mathbf{a}}(O)$ for its law --- the distribution of that value
  over the draw of $\rho$.

  The first reading is an equality of laws. Definition~\ref{def:d-probing} says of
  a $d$-probing secure circuit that
  \[
    \mathcal{L}_{\mathbf{a}}(O) = \mathcal{L}_{\mathbf{b}}(O)
    \qquad \text{for all } \mathbf{a}, \mathbf{b} \in \{0,1\}^{\mathcal{S}}
    \text{ and every } O \text{ with } |O| \le d ,
  \]
  so the law of a probe is one and the same whatever the secrets are, hence we cannot
  deduce information about the secrets by looking at the values on the probe.

  The second reading is \emph{simulability}, and it is the one that names
  non-interference. A probe $O$ is non-interfering when some randomised algorithm
  $\mathsf{Sim}(O)$, given the probe and the public inputs but never the secrets,
  has output law equal to $\mathcal{L}_{\mathbf{a}}(O)$ for every $\mathbf{a}$. Such
  a probe shows the adversary nothing it could not have produced on its own without
  the secret. Simulability and
  equality of laws are the same property. If the laws agree, their common value is a
  distribution the simulator samples outright, and conversely a single simulator
  that matches every $\mathbf{a}$ forces all the $\mathcal{L}_{\mathbf{a}}(O)$ to
  coincide with its output, hence with one another.

  The third reading is a game, and it is the sharpest picture of what security
  entails for the adversary. The adversary names a probe $O$ of size at most $d$ together
  with two secret assignments $\mathbf{a}_0$ and $\mathbf{a}_1$. A challenger draws
  a hidden bit $c \in \{0,1\}$ and fresh randomness $\rho$, runs the circuit under
  $\mathbf{a}_c$, and hands back the transcript $O(\rho; \mathbf{a}_c)$. The
  adversary guesses $c$, and its advantage $\big|\Pr[\text{guess} = c] -
  \tfrac{1}{2}\big|$ is at most the statistical distance
  \[
    \tfrac{1}{2} \sum_{o} \big| \mathcal{L}_{\mathbf{a}_0}(O)(o)
      - \mathcal{L}_{\mathbf{a}_1}(O)(o) \big|
  \]
  between the two laws, with equality for the optimal guess. The circuit is
  $d$-probing secure exactly when this distance is zero for every $O$,
  $\mathbf{a}_0$, and $\mathbf{a}_1$, so that the two secret worlds are perfectly
  indistinguishable through any $d$ wires. We call this
  \emph{secret-indistinguishability}.

  The three are one condition seen from three sides, equal laws, a secret-free
  simulator, and a game no adversary wins.
  % To show that a probe is non-interfering, it is enough to
  % rewrite its wires into a form in which the secrets have cancelled, which is the
  % program logic Barthe et al.\ introduce and the engine of
  % Chapter~\ref{chap:design} mechanises.

  % I figured, because we do not treat composability in the thesis. We don't really
  % need to introduce it in the preliminaries chapter.
  % \subsection{Threshold and Strong Non-Interference}
  % \label{sec:sni}

  % Probing security, unfourtanetly, does not compose. A secure gadget can
  % leak, once its output feeds another, because probes in the second gadget might see
  % combinations the first one never accounted for. Barthe et al.
  % \parencite{barthe2016strong} isolate two stronger notions that do compose. A
  % gadget is \emph{non-interfering} (NI) when any $d$ probes on it can be simulated
  % from at most $d$ shares of each input, and \emph{strongly non-interfering} (SNI)
  % when the same holds under a sharper budget that counts only the probes on
  % internal wires, not those on the outputs. SNI is what lets one assemble a large
  % masked circuit from small verified gadgets while preserving the order. These
  % notions organise much of the field, though the thesis stays with plain probing
  % security and does not pursue composition, so we recall them mainly to place our
  % object among its neighbours.

\section{Randomness Substitution}
\label{sec:randomness-substitution}

This section isolates, as a single reusable fact, the probabilistic argument
our checker automates. Fix a circuit as in Section~\ref{sec:bool-circuits},
with random inputs $\mathcal{R}$, and let a \emph{randomness assignment}, or
\emph{tape}, be a function $\rho : \mathcal{R} \to \{0,1\}$. The secrets and public
inputs are held fixed throughout, so every wire $w$ denotes a function $w(\rho)$ of
$\rho$ alone. A \emph{probe} is an observation set of Section~\ref{sec:d-probing},
which we list as a tuple $\mathbf{w}$ of wires; its order does not affect the joint
distribution. That section phrases $d$-probing security as the demand that the
joint distribution of a probe of size at most $d$, taken over the uniformly
distributed $\rho$, not move as the secrets vary.

The whole argument rests on one line of $\mathbb{F}_2$ arithmetic: for any fixed
bit $c$, the translation/shear $b \mapsto c \oplus b$ is a \textbf{bijection} of $\{0,1\}$
that preserves the uniform distribution (i.e. it is its own inverse and merely swaps
the two outcomes). Everything below is this fact, applied to a single coordinate
of $\rho$ with every other coordinate fixed.

\begin{definition}[Randomness substitution]
\label{def:substitution}
Let $r \in \mathcal{R}$ and let $e$ be a wire with $r \notin \mathrm{vars}(e)$.
The \emph{substitution} $\sigma = (r \mapsto e \oplus r)$ acts on a wire $w$ by
replacing every occurrence of the input $r$ with the wire $e \oplus r$; we write
the resulting wire as $w\sigma$. The same $\sigma$ acts on randomness assignments,
through the map $\widehat{\sigma} : \{0,1\}^{\mathcal{R}} \to
\{0,1\}^{\mathcal{R}}$ whose image $\rho' = \widehat{\sigma}(\rho)$ shifts the
$r$-coordinate by $e(\rho)$ and leaves every other coordinate untouched:
\[
  \rho'(r) = e(\rho) \oplus \rho(r), \qquad
  \rho'(r') = \rho(r') \quad \text{for } r' \neq r .
\]
\end{definition}

Geometrically, $\widehat{\sigma}$ is a \emph{shear} of the cube
$\{0,1\}^{\mathcal{R}}$: it fixes every coordinate but $r$ and offsets that one
by $e(\rho)$. Restricted to a
coordinate it is
exactly the translation $\rho(r) \mapsto e(\rho) \oplus \rho(r)$, so it preserves the measure
by translation invariance.
We call any such map, one coordinate translated by a context
depending on the others, and everything else left intact, a \emph{shear}, and use
the word throughout.

We pair the two actions under one name with the lemma below. The
syntactic rewrite $w \mapsto w\sigma$ and the reparametrisation
$\rho \mapsto \widehat{\sigma}(\rho)$ are one operation seen from two sides, and
$\widehat{\sigma}$ does not disturb the distribution.

\begin{lemma}[Substitution]
\label{lem:substitution}
For a substitution $\sigma = (r \mapsto e \oplus r)$ as in
Definition~\ref{def:substitution}, the map $\widehat{\sigma}$ is a
measure-preserving involution of the uniform distribution on
$\{0,1\}^{\mathcal{R}}$, and for every wire $w$,
\[
  (w\sigma)(\rho) \;=\; w\big(\widehat{\sigma}(\rho)\big)
  \qquad \text{for every randomness assignment } \rho .
\]
\end{lemma}

\begin{proof}
Fix every coordinate of $\rho$ other than $r$. Since $r \notin \mathrm{vars}(e)$,
the value $e(\rho)$ depends only on these fixed coordinates, so it is constant
as $\rho(r)$ ranges over $\{0,1\}$. On this fibre $\widehat{\sigma}$ acts by the
translation $\rho(r) \mapsto e(\rho) \oplus \rho(r)$, which is measure-preserving.
Hence $\widehat{\sigma}$ is a measure-preserving involution of $\{0,1\}^{\mathcal{R}}$
preserving the uniform measure coordinatewise, and so preserving the product
measure.

To prove the stated identity, do a structural induction on $w$. For $w = r$ both sides equal
$e(\rho) \oplus \rho(r)$. For an input $w \neq r$ both sides equal $w(\rho)$,
since $\widehat{\sigma}$ fixes every coordinate but $r$, they are also equal. For a gate
$w_i = w_j \oplus w_k$ or $w_i = w_j \odot w_k$, just apply the hypothesis to $w_j$
and $w_k$ and use that $\oplus$ and $\odot$ are computed pointwise from their
operands' values.
\end{proof}

\begin{corollary}[Single-occurrence discharge]
\label{cor:single-occurrence-discharge}
Suppose $r$ occurs exactly once among a probe $\mathbf{w}$ and the wires it
depends on, as an operand of a gate $x = e \oplus r$ with
$r \notin \mathrm{vars}(e)$, and let $\sigma = (r \mapsto e \oplus r)$. Then
$\mathbf{w}\sigma$ is identically distributed to $\mathbf{w}$ over $\rho$
uniform, and every occurrence of $x$ in $\mathbf{w}\sigma$ has collapsed to the
bare input $r$.
\end{corollary}

\begin{proof}
Distributional equality is Lemma~\ref{lem:substitution} applied coordinatewise
to $\mathbf{w}$, using that $\widehat{\sigma}$ preserves the uniform
distribution on $\{0,1\}^{\mathcal{R}}$. For the second claim,
$x\sigma = e \oplus (e \oplus r) = r$ by $\mathbb{F}_2$-cancellation; since $r$
occurs nowhere else by hypothesis, no other part of $\mathbf{w}$ is touched.
\end{proof}

\begin{remark}
Rewriting a probe by a substitution --- whether in the general form of
Lemma~\ref{lem:substitution} or the single-occurrence case of
Corollary~\ref{cor:single-occurrence-discharge} --- leaves its distribution
untouched. That is the key observation that derives the procedure
to decide the security of a probe, by one
substitution at a time, taking it to a form with no secret in sight. The
single-occurrence case is the one the \texttt{maskVerif} tool
\parencite{10.1007/978-3-030-29959-0_15} calls \emph{optimistic sampling}. A
derivation is nothing more than such substitutions chained together; how that
chain is searched for is taken up in the later chapters.
\end{remark}

\section{Computational Hardness Preview}
\label{sec:hardness-preview}

Definition~\ref{def:d-probing} asks, for each observation set, whether a
distribution stays fixed as the secrets vary, and read literally this is a
counting question, beacuse the measures we work with are defined on discrete spaces endowed with the counting measure.
The law of a probe assigns to each possible observation the
number of randomness assignments $\rho$ that produce it, divided by
$2^{|\mathcal{R}|}$, and independence of the secrets asks that these counts agree
across secret settings. Chapter~\ref{chap:couplings} makes the difficulty precise,
showing that the associated counting problem is \#P-complete, \#P being the class
Valiant introduced for counting the accepting computations of a nondeterministic
machine \parencite{valiant1979complexity}.

The verification question of $d$-probing security, however, is a \emph{decision} problem rather than
a counting one, so it cannot be a member of \#P. Deciding whether one probe is
independent of the secrets compares two counts and asks whether they are equal, and
the class of decision problems of this shape, i.e. whether a nondeterministic machine
has exactly a prescribed number of accepting paths, is $\mathrm{C}_{=}\mathrm{P}$,
the exact-counting counterpart of \#P studied by Wagner
\parencite{wagner1986complexity}. Its canonical complete problem, deciding whether
a Boolean formula has exactly a given number of satisfying assignments, is the
equality-of-counts question in its essence, and probing security is that same
question, but posed in the language of masked circuits. Locating the problem in
$\mathrm{C}_{=}\mathrm{P}$ says plainly that an exact decision procedure is not
expected to be efficient at all, and it is the reason this thesis does not count.
It certifies security constructively instead, through the rewriting of
Section~\ref{sec:randomness-substitution}, giving a checker that is sound but
necessarily incomplete.
