\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Related Work}
\label{chap:related}
% ------------------------------------------------------------
Verifying that a masked circuit leaks nothing has been approached from several
sides.
The question is always the one of Chapter~\ref{chap:preliminaries}, asking whether the joint
distribution of a set of wires vary with the secrets, but one may resolve it by
rewriting an expression, by measuring a spectrum, by counting satisfying
assignments, or by building the distributions explicitly. This chapter surveys those
approaches, and places our work among them.

\section{Language-Based Verification} % main inspiration
\label{sec:language-based}

The approach we build on views a masked circuit as a probabilistic program and
treats a probe's leakage symbolically. Rather than estimating a distribution, it
manipulates the expressions a set of wires compute until secrets have
visibly cancelled. The line begins with \parencite{barthe2015verified},
and their framing of secret-independence as a property that can be established by
distribution-preserving rewriting, is the largest influence of this work.

  \subsection{Verified Proofs of Higher-Order Masking}
  \label{sec:related-barthe}

  Barthe et al.\ \parencite{barthe2015verified} gave a machine-checked
  proof of higher-order masking, built inside the EasyCrypt proof assistant and
  its probabilistic relational Hoare logic. Their checker decides whether a set of % add easycrypt ref
  intermediate wires is independent of the secret by rewriting. The central
  method is called optimistic sampling, where a random $r$ occurring once, as
  $e \oplus r$, is replaced by a fresh uniform variable, since $e \oplus r$ is
  itself uniform when $r$ does not appear in $e$. When substitutions alone
  fail to cancel all secrets, the expression is put into algebraic normal form so that
  products collapse, and independence can be declared if no secret variable occurs in that form.
  This is, in miniature, the rewrite engine of our Component~I
  (Section~\ref{sec:rewrite-engine}). The optimistic sampling substitution is
  exactly the single-occurrence rule we inherit. However, where they go to ANF,
  we instead factor the expression (Section~\ref{sec:factoring}), which can reach
  products that normalisation miss.

  \subsection{Large Sets and the Exploration Algorithm}
  \label{sec:related-largesets}

  Deciding one set is not the only difficulty; indeed, there are
  $\sum_{i \le d} \binom{|\mathcal{W}|}{i}$ of them. The language-based tools escape
  this enumeration with a large-set argument that the \texttt{maskVerif} tool
  \parencite{10.1007/978-3-030-29959-0_15} made systematic. Secret-independence
  passes to subsets, so a proof about a large set of wires discharges every
  one of its subsets at once. So, the search tries to verify maximal sets and, on failure,
  splits and recurses. Our Component~II
  (Section~\ref{sec:subset-space}) is this exploration, and our contribution there
  is to make the split's exhaustiveness a neatly proved identity. The set-level
  Vandermonde partition of Section~\ref{sec:vandermonde} certifies that no probe
  slips through such a recursion.

  \subsection{Witness Replay and Derivation Reuse}
  \label{sec:related-replay}

  A certified set leaves behind a derivation, which is the sequence of rewrites that
  discharged it, and that derivation can be reused. \texttt{maskVerif}
  \parencite{10.1007/978-3-030-29959-0_15} enlarges a verified set by replaying its
  derivation against neighbouring positions and keeping those that are also carried to a
  secret-free form after the same sequence of rewrites.
  So, one certification serves many wires. This is the direct
  ancestor of our Component~III (Section~\ref{sec:replay}). The reuse is imperfect
  for us, because our engine factors and factoring is not a trivially replayable step. Re-running
  a recorded derivation on a fresh wire misses the cancellations a factoring might have
  made possible. We therefore read the derivation as a
  measure-preserving coupling of the randomness (Chapter~\ref{chap:couplings}), and
  we test each candidate against it with the exact criterion of
  Section~\ref{sec:coupling-extension}, which admits every wire replays also would.

  \subsection{maskVerif and Physical Defaults}
  \label{sec:related-maskverif}

  The value-based model, in which a probed wire leaks solely its logical value, is
  very optimistic about issues in hardware. A real gate \emph{glitches}, briefly computing on unsettled
  inputs, and a register transition may leak two successive values.
  \texttt{maskVerif} \parencite{10.1007/978-3-030-29959-0_15} captures
  these physical defaults, verifying masking in a glitch-extended probing
  model where a probe on a glitching gate carries information about all of that gate's stable
  inputs.
  We stay in the value-based model and do not treat glitches
  (Chapter~\ref{chap:preliminaries}), so those stronger models which do, bound the guarantee
  our checker delivers.

\section{Fourier-Analytic Verification} % Bloem
\label{sec:fourier-analytic}

A different approach does not rewrite at all. Observe that a set of wires is
independent of the secret exactly when certain Fourier coefficients of the Boolean
function they compute vanish, and one can verify masking by checking that spectral
condition.

  \subsection{The Xiao\textendash Massey Criterion and Approximations}
  \label{sec:related-xiaomassey}

  The spectral implications of independence is widely known. Xiao and Massey
  \parencite{xiao1988spectral} characterised the correlation-immune Boolean
  functions as those whose Walsh spectrum vanishes at every nonzero point of
  Hamming weight at most $t$. So, a function is $t$-th order correlation-immune, that
  is, statistically independent of any $t$ of its inputs, precisely under this
  condition. Bloem et al.\ \parencite{bloem2018formal} carry this into masking
  verification. Their tool, \texttt{REBECCA}, propagates a conservative correlation set
  for each wire through the circuit. This set consists of monomials that might have a non-zero
  Fourier coefficient (indicating potential correlation), and the tool reports a
  leak when such a monomial mixes a secret with no fresh masking variable to blind it.
  Tracking which monomials can appear, rather than their exact coefficients, is the
  over-approximation that keeps the computation feasible, at the cost of the occasional
  false positives.

  \subsection{SAT-Based Realization and Its Cost}
  \label{sec:related-sat}

  The correlation-immunity condition is discharged to a SAT solver. \texttt{REBECCA}
  \parencite{bloem2018formal} reduces the search for a secret-dependent, unblinded
  correlation to a family of SAT queries whose number is polynomial in the count of
  gates and wires, and it runs in the glitch-extended model, making it one of the
  first tools to verify hardware masking under physical defaults. The correlation
  sets get increasingly convoluted with the masking order, so the propagation and
  the satisfiability search both climb as the order grows, and the over-approximation
  has to be loosened or tightened to trade precision against effort.

\section{Other Approaches}
\label{sec:related-other}

Beyond rewriting and spectra, two further approaches are worth mentioning. One decides
independence by educated counting, the other sidesteps the per-circuit question
by reasoning about compositions of smaller gadgets.

  \subsection{Model Counting and SMT}
  \label{sec:related-modelcounting}

  Independence is a statement about a distribution, so one can check it by encoding
  the distribution and asking a solver. Eldib, Wang, and Schaumont
  \parencite{eldib2014formal} verify software masking by expressing
  secret-independence as a satisfiability question and discharging it to an SMT
  solver. Gao et al.\ \parencite{gao2019verifying} makes the question quantitative,
  asking how far a wire is from independence through a model-counting notion of
  masking strength and scaling it with a refinement loop. \texttt{SILVER}
  \parencite{knichel2020silver} goes fully exact, constructing the wires' joint
  distribution with reduced-ordered binary decision diagrams and testing
  statistical independence on it directly. These tools are exact where ours is
  incomplete, and the reason we do not follow them is the hardness of
  Chapter~\ref{chap:couplings}. Computing or comparing these distributions is
  $\#\mathrm{P}$- and $\mathrm{C}_{=}\mathrm{P}$-hard, so the exact route is taken
  with a heavy cost.

  \subsection{Compositional Reasoning and Relabeling}
  \label{sec:related-composition}

  Probing security does not trivially compose. A gadget which is secure in isolation can leak once
  wired into another secure one. A body of work isolates stronger notions that allow compositions. Barthe et
  al.\ \parencite{barthe2016strong} introduce strong non-interference, which handle
  output and internal probes separately and composes by construction, and Cassiers
  and Standaert \parencite{cassiers2020trivially} refine this to a probe-isolating
  non-interference, under which gadgets compose nicely. Belaïd, Goudarzi, and
  Rivain \parencite{belaid2018tight} verify tight probing security directly,
  characterising when a masked circuit is secure with the least refreshing. These
  notions allow the assembly of large masked circuits from smaller components. We stay within
  plain probing security of a single circuit (Chapter~\ref{chap:preliminaries}) and do
  not pursue composition, so they are related to our work but do not overlap it.

\section{Position of This Work}
\label{sec:related-position}

This thesis sits in the language-based lineage of Barthe et al.\
\parencite{barthe2015verified, 10.1007/978-3-030-29959-0_15}. As they do, we decide
a probe by rewriting it toward a secret-free form, we escape the subset enumeration
by exploring large sets, and we reuse a certified derivation to enlarge a safe set.
Three things set our treatment apart. First, we view a successful discharge as an
explicit measure-preserving coupling of the randomness
(Chapter~\ref{chap:couplings}) and admit neighbouring wires by an exact test
through it. Second, we prove the
exhaustiveness of the set-splitting, turning the traversal's soundness into the
Vandermonde identity of Section~\ref{sec:vandermonde} rather than an informal % bu kismi suha hocaya bi danis
argument. Third, we add multiplicative factoring (Section~\ref{sec:factoring}) to
reach randoms buried inside products.
Unlike the Fourier and BDD
tools we stay value-based and do not model glitches, and unlike the compositional
notions we verify a whole circuit rather than assembling it from verified parts.
Against the exact solvers and decision-diagram engines we decline to build the
distribution at all, since Chapter~\ref{chap:couplings} shows that to be
intractable, and we certify constructively instead, which is sound but incomplete.
