\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Couplings and Probing Security}
\label{chap:couplings}
% ------------------------------------------------------------
\section{The Complexity of Deciding Secret-Independence}
\label{sec:hardness}

Definition~\ref{def:d-probing} phrases security as the condition that the law of a
probe stay fixed as the secrets vary, and Section~\ref{sec:hardness-preview}
already mentioned that this demand is actually a counting question. In this section we make
this claim precise. Underneath a single probe sit two problems, computing its law
and deciding whether varying the secret move it, and we determine each one's hardness. Computing the law
is a $\#\mathrm{P}$-complete count, and the decision problem is a
$\mathrm{C}_{=}\mathrm{P}$-complete comparison of such counts. This means
that verifying even one probe by counting is complete for classes that
sit above the entire polynomial hierarchy, which is the reason the thesis certifies
security constructively without explicit counting.

Throughout, fix a circuit over $\mathbb{F}_2$ with secret inputs $\mathcal{S}$,
random inputs $\mathcal{R}$, and public inputs $\mathcal{P}$, as usual, together with a probe
$O = \{w_1, \dots, w_k\}$ of size $k \le d$. Under a secret assignment
$\mathbf{a} \in \{0,1\}^{\mathcal{S}}$ and a tape
$\rho \in \{0,1\}^{\mathcal{R}}$, the probe takes the joint value
$O(\rho; \mathbf{a}) \in \{0,1\}^{k}$, with the publics held fixed.

  \subsection{From Independence to Distribution Counting}
  \label{sec:indep-to-counting}

  The law of a probe is assembled from integer counts. For an observation
  $o \in \{0,1\}^{k}$ write
  \[
    N_{\mathbf{a}}(o) \;=\; \big|\{\, \rho \in \{0,1\}^{\mathcal{R}} : O(\rho; \mathbf{a}) = o \,\}\big| ,
  \]
  the number of tapes on which the probe reads $o$ under the secret $\mathbf{a}$.
  Since $\rho$ is uniform, the law is
  $\mathcal{L}_{\mathbf{a}}(O)(o) = N_{\mathbf{a}}(o) / 2^{|\mathcal{R}|}$, so the
  law is nothing but a histogram of these counts. Two laws coincide
  exactly when their counts do,
  \[
    \mathcal{L}_{\mathbf{a}}(O) = \mathcal{L}_{\mathbf{b}}(O)
    \quad\Longleftrightarrow\quad
    N_{\mathbf{a}}(o) = N_{\mathbf{b}}(o) \ \text{ for every } o \in \{0,1\}^{k} .
  \]
  Secret-independence (Definition~\ref{def:d-probing}) is this equality quantified
  over all pairs of secrets. So what looks like a statement about distributions is
  actually a statement about integers, that certain counts of satisfying tapes agree.
  This splits into two questions of different character. One asks for the count itself;
  the other asks only whether two counts are equal, a decision. We treat them in turn.

  \begin{definition}[The counting and distinguishing problems]
  \label{def:count-distinguish}
  \textsc{Count} takes a circuit, a probe $O$, a secret $\mathbf{a}$, and an
  observation $o$, and returns the integer $N_{\mathbf{a}}(o)$.
  \textsc{Distinguish} takes a circuit, a probe $O$, and two secret assignments
  $\mathbf{a}, \mathbf{b}$, and accepts exactly when
  $\mathcal{L}_{\mathbf{a}}(O) = \mathcal{L}_{\mathbf{b}}(O)$.
  \end{definition}

  \subsection{Counting the Law is \NoCaseChange{\#P}-Complete}
  \label{sec:sharpp}

  Recall that $\#\mathrm{P}$ is Valiant's class of functions counting the accepting
  computations of a nondeterministic polynomial-time machine, equivalently it is the
  enumeration of the satisfying assignments of a polynomial-size Boolean predicate
  \parencite{valiant1979complexity}.

  \begin{proposition}[Counting the law]
  \label{prop:count-sharpp}
  \textsc{Count} is $\#\mathrm{P}$-complete.
  \end{proposition}

  \begin{proof}
  \emph{Membership.} Fix $\mathbf{a}$, the publics, and $o$. The predicate
  $\rho \mapsto [\, O(\rho; \mathbf{a}) = o \,]$ evaluates the circuit on the tape
  $\rho$ and compares $k$ output bits against $o$, all in time polynomial in the
  circuit size. Hence $N_{\mathbf{a}}(o)$ counts the tapes satisfying a
  polynomial-time predicate, so it is a $\#\mathrm{P}$ function.

  \emph{Hardness.} We reduce from $\#\textsc{SAT}$, the count of satisfying
  assignments of a Boolean formula, which is $\#\mathrm{P}$-complete
  \parencite{valiant1979complexity}. Given $\varphi(x_1, \dots, x_m)$, build the
  circuit that reads $x_1, \dots, x_m$ as random inputs, has no secrets, and
  computes $\varphi$ on a single wire $w$, realising the connectives of $\varphi$
  through $\oplus$ and $\odot$, which is possible because the two are universal over
  $\mathbb{F}_2$ (Section~\ref{sec:bool-circuits}). The probe $\{w\}$ then has
  $N(1) = |\{ x : \varphi(x) = 1 \}| = \#\varphi$, so any procedure for
  \textsc{Count} computes $\#\varphi$. The reduction is polynomial, hence
  \textsc{Count} is $\#\mathrm{P}$-hard.
  \end{proof}

  Notice that the hardness result uses no secret at all, so computing a probe's law is already
  intractable on the plainest circuits. That count is the object every exact method
  must, explicity or implicitly, realize.

  % The class wraps every \subsection title in titlecaps' \titlecap, which loops on
  % a math subscript; \oldsubsection bypasses that wrapper for this one heading.
  \oldsubsection{Deciding Independence is \texorpdfstring{\ensuremath{\mathrm{C}_{=}\mathrm{P}}}{C=P}-Complete}
  \label{sec:ceqp}

  \textsc{Distinguish} never asks for a count outright. It asks only whether two
  counts are equal, and that equality has a class of its own. Following Wagner,
  $\mathrm{C}_{=}\mathrm{P}$ is the class of languages of the form ``does a
  nondeterministic polynomial-time machine have exactly a given number of
  accepting paths'' \parencite{wagner1986complexity}. Another view comes
  from gap functions. Let $\mathrm{GapP}$ be the closure of $\#\mathrm{P}$ under
  subtraction, so the differences $f - g$ of two counting functions
  \parencite{fenner1994gap}. Then $\mathrm{C}_{=}\mathrm{P}$ is exactly the class of
  vanishing sets of gap functions,
  \[
    \big\{\, x \mid h(x) = 0 \,\big\}, \qquad h \in \mathrm{GapP} .
  \]
  Its canonical complete problem is \textsc{ExactSat}, deciding whether a formula
  has exactly a given number of satisfying assignments, equivalently one can also
  ask whether two formulas have the same number of satisfying assigmnets
  \parencite{wagner1986complexity}.

  \begin{theorem}[Deciding independence]
  \label{thm:distinguish-ceqp}
  \textsc{Distinguish} is $\mathrm{C}_{=}\mathrm{P}$-complete. Consequently deciding
  whether a probe is secret-independent is $\mathrm{C}_{=}\mathrm{P}$-hard.
  \end{theorem}

  \begin{proof}
  \emph{Membership.} Put
  \[
    f(C, O, \mathbf{a}, \mathbf{b}) \;=\; \sum_{o \in \{0,1\}^{k}}
      \big(N_{\mathbf{a}}(o) - N_{\mathbf{b}}(o)\big)^2 .
  \]
  Each $N_{\mathbf{a}}$ is a $\#\mathrm{P}$ function, and $\mathrm{GapP}$ is closed under
  subtraction and multiplication \parencite{fenner1994gap}. The sum ranges over
  $2^{k} \le 2^{d}$ observations, which is a constant since the order $d$ is fixed, thus $f$ is
  a constant combination of gap functions and is itself in $\mathrm{GapP}$. Remember that a sum of
  squares vanishes exactly when every term does, hence
  \[
    \mathcal{L}_{\mathbf{a}}(O) = \mathcal{L}_{\mathbf{b}}(O)
    \iff N_{\mathbf{a}}(o) = N_{\mathbf{b}}(o) \ \forall o
    \iff f(C, O, \mathbf{a}, \mathbf{b}) = 0 ,
  \]
  and \textsc{Distinguish} is the vanishing set of the gap function $f$, so it lies in
  $\mathrm{C}_{=}\mathrm{P}$.

  \emph{Hardness.} We reduce from the equality form of \textsc{ExactSat}: given two
  formulas $\varphi_0, \varphi_1$ over $x_1, \dots, x_m$, decide whether
  $\#\varphi_0 = \#\varphi_1$, which is $\mathrm{C}_{=}\mathrm{P}$-complete
  \parencite{wagner1986complexity}. Build a circuit with a single secret bit $s$,
  the $x_i$ as randoms, and one wire
  \[
    w \;=\; \varphi_0 \,\oplus\, \big(s \odot (\varphi_0 \oplus \varphi_1)\big) ,
  \]
  again realising the formulas over $\oplus$ and $\odot$. Then $w$ reads
  $\varphi_0$ when $s = 0$ and $\varphi_1$ when $s = 1$, so
  $N_{s=0}(1) = \#\varphi_0$ and $N_{s=1}(1) = \#\varphi_1$. The one-bit probe
  $\{w\}$ satisfies $\mathcal{L}_{0} = \mathcal{L}_{1}$ if and only if
  $\#\varphi_0 = \#\varphi_1$, so \textsc{Distinguish} on this circuit with
  $\mathbf{a} = 0$, $\mathbf{b} = 1$ decides the equality. The construction is
  polynomial, so \textsc{Distinguish} is $\mathrm{C}_{=}\mathrm{P}$-hard. One secret
  bit and one probed wire already suffice, so no feature of the masking is what
  makes the problem hard.
  \end{proof}

  \subsection{What the Placements Mean for Verification}
  \label{sec:hardness-consequences}

  The two results are closely related. The law's entries are $\#\mathrm{P}$
  counts, and the security question is whether two of them coincide, which is the
  defining shape of the $\mathrm{C}_{=}\mathrm{P}$ class. The classes fall in the expected
  places. Every $\mathrm{C}_{=}\mathrm{P}$ query is settled by two calls to a
  counting oracle, followed by a subtraction, and a test against zero, so
  $\mathrm{C}_{=}\mathrm{P} \subseteq \mathrm{P}^{\#\mathrm{P}}$. Which means
  that the decision never costs more than the counting.

  Another observation is the following: a formula is unsatisfiable exactly when its count is
  zero (as an instance of exact counting) so $\mathrm{coNP} \subseteq
  \mathrm{C}_{=}\mathrm{P}$. Hence \textsc{Distinguish} is already
  $\mathrm{coNP}$-hard, and no polynomial-time exact decision exists unless
  $\mathrm{P} = \mathrm{NP}$. Further showing us that deciding
  one probe exactly is very much intractable.

  The full verification task only adds on this. $d$-probing security asks \textsc{Distinguish}
  to answer ``equal'' for every one of the $\sum_{i \le d} \binom{|\mathcal{W}|}{i}$
  probes and every pair of secrets. A universal quantifier over exponentially many
  instances of a $\mathrm{C}_{=}\mathrm{P}$-complete question.

  So now we should be fairly certain that computing or
  equating these counts is out of reach, so we do not compute them. We certify security
  constructively instead, by exhibiting for a probe a measure-preserving relabelling
  of the randomness under which its secrets visibly cancel. Such a witness proves independence without ever
  counting, though at the price of completeness, since a probe can be independent for
  reasons no relabelling of our kind captures.

\section{Couplings of Probabilistic Programs}
\todo[inline]{Introduce couplings as a proof technique for distributional equality between probabilistic programs.}
  
  \subsection{Coupling Two Runs Under Different Secrets}
  \todo[inline]{Define a coupling between the run of a probe under one secret assignment and its run under another.}
  
  \subsection{Coupling Functions: Randoms to Randoms} % measure-preserving bijections on randomness
  \todo[inline]{Formalize coupling functions as measure-preserving bijections on the randomness domain.}

  \section{Probing Security via Couplings}
  \todo[inline]{Relate $d$-probing security to the existence of a coupling for every probe of size $d$.}
  
  \subsection{\NoCaseChange{$d$}-Probing Security Implies a (Family of) Coupling Functions}
  \todo[inline]{Prove that $d$-probing security is equivalent to the existence of a family of coupling functions, one per pair of secret assignments.}
  
  \subsection{Secret-Indistinguishability}
  \todo[inline]{Connect the coupling-existence characterization back to the standard secret-indistinguishability definition of probing security.}

\section{Substitutions as a Coupling Construction}
\todo[inline]{Explain how the randomness-substitution rewrites of Section~\ref{sec:randomness-substitution} construct an explicit coupling.}
  
  \subsection{The Derivation of an Explicit Coupling Witness}
  \todo[inline]{Show how a derivation (sequence of substitutions) yields an explicit coupling witness for a probe.}
