\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Implementation of Coppula}
\label{chap:implementation}
% ------------------------------------------------------------
\section{Expression Representation}
\label{sec:expr-representation}

The rewrite engine of Component~I (Section~\ref{sec:rewrite-engine}) works
entirely on the expressions of a probe. It looks for a random in an additive
context, substitutes it by its context, and asks whether any secret is still in
sight. Every one of these moves is an operation on the \emph{form} of the
expression, so the structure that holds that form is what decides how cheap the
engine can be. This section describes that structure. It is a hash-consed
$n$-ary DAG over $\mathbb{F}_2$, wrapped in a layer of smart constructors that
keep every expression canonical, and extended with a multiplicative factoring
pass that reshapes products so the engine can reach randoms otherwise buried
inside them.

One point of clarification before we dive in. The circuit of
Definition~\ref{def:circuit} is strictly $2$-ary, and that is deliberate. Each
$2$-input gate output is a distinct observable, so collapsing gates would hide
intermediate wires the adversary is allowed to probe
(Section~\ref{sec:d-probing}). The representation here is not the circuit, and this important.
It is how we store the \emph{value} of a wire for the algebra, and there flattening a
chain of same-typed gates into one $n$-ary node loses nothing, since only the
wire's function matters and not its gate-by-gate history. So the two structures
(one faitfhul to the topology and one keeping the expressions) live side
by side. The $2$-ary netlist stays the authority on what may be probed, and the
DAG is the authority on what each probed wire computes.

  \subsection{A Hash-Consed \NoCaseChange{$n$}-ary Circuit DAG}
  \label{sec:hash-consed-dag}

  An expression is a node of one of four kinds. It is a \emph{leaf} carrying an
  input variable (secret, random, or public), a \emph{constant} $0$ or $1$, an
  $n$-ary sum $\oplus$ over a list of operand nodes, or an $n$-ary product
  $\odot$ over a list of operand nodes. A wire's value is the node its defining
  sub-circuit builds, and the whole circuit is a single DAG in which wires that
  share sub-structure share nodes.

  Sharing is the first reason for the design, and it is not incidental. A masked
  gadget reuses the same handful of masks and shares across many gates, so the
  naive syntax tree of a probe would carry the same subexpression over and over again.
  We store each distinct node once. A table, keyed by a node's kind together with
  its operand list, which maps a structure to the identifier of the node already built
  for it; this is \emph{hash-consing}. Constructing a node is then a lookup in
  that table. On a hit we return the existing identifier, and on a miss we
  allocate a fresh one and record it. Two expressions that are structurally the
  same are therefore the same node, physically, however many times they are
  written.

  The second reason is equality. Because interning makes structural identity
  coincide with identifier identity, deciding whether two expressions are equal
  is a comparison of two integers, with no traversal at all. The rewrite engine
  leans on this everywhere. Recognising that a substitution has made two operands
  coincide so that they cancel, checking whether a rewritten probe matches one
  already seen, and, in Component~III, testing a candidate wire's coupling image
  against a secret-free target, are each an $O(1)$ test on identifiers rather than
  a walk over two trees. We should note that this fast equality is only as
  meaningful as the expressions are canonical: hash-consing alone shares
  expressions written identically, and it is the next subsection that makes
  identical identifiers mean equal values. We also keep the identifiers of the
  random and secret leaves in their own lists, so iterating over
  $\mathrm{vars}(\mathbf{w})$, or scanning for a surviving secret, never has to
  walk the entire DAG.

  \subsection{Canonicalization: Flattening and Cancellation}
  \label{sec:canonicalization}

  Hash-consing shares expressions that are written identically, which on its own
  is weak. In $\mathbb{F}_2$ the sums $a \oplus b$ and $b \oplus a$ are equal,
  yet written in that order they are two different nodes. We close this gap with a
  layer of \emph{smart constructors}. Nothing builds a sum or a product node
  directly; every sum is made through one constructor and every product through
  another, and these put the operands into a canonical form before interning.
  Two expressions with the same value then reach the same canonical node, so the
  fast equality of Section~\ref{sec:hash-consed-dag} becomes equality of value
  and not merely of syntax, at least for the identities we fold in, which are
  exactly the ones the engine needs.

  Three normalisations do the work. The first is \emph{flattening}.
  Associativity lets a sum of sums be one flat sum, and likewise a product of
  products be one flat product. This is where the $n$-ary shape earns itself, as
  a chain of $2$-ary XOR gates from the netlist collapses into a single sum node
  holding all the leaves, and any nesting is quotiented away. The second is
  \emph{ordering}. Commutativity lets us keep each operand list sorted by
  identifier, so the order of writing is irrelevant to identity and $a \oplus b$
  and $b \oplus a$ intern to one node. The third folds in the $\mathbb{F}_2$
  identities the engine's cancellations depend on. In a sum, the constant $0$ is
  dropped as the additive identity, and a repeated operand is deleted in pairs,
  since $x \oplus x = 0$.
  In a product, the constant $1$ is dropped as the multiplicative identity, the
  constant $0$ collapses the whole product to $0$ by annihilation, and a repeated
  operand is kept only once, since $x \odot x = x$ by idempotence
  (Section~\ref{sec:field-f2}). Degenerate arities settle the same way. An empty
  sum is $0$, an empty product is $1$, and a one-operand sum or product is that
  operand.

  The payoff is that the one algebraic move the whole rewrite method turns on
  happens for free. When the engine substitutes a single-occurrence random $r$ by
  its context, replacing the parent $x = e \oplus r$, the cancellation
  $e \oplus (e \oplus r) = r$ that Corollary~\ref{cor:single-occurrence-discharge}
  promises is not a step the engine performs. It re-interns the rewritten
  expression through the sum constructor, the two copies of $e$ meet in the sorted
  operand list and annihilate, and the canonical result is already the collapsed
  form. The idempotent and annihilating collapses that a general substitution
  triggers happen the same way. The engine proposes rewrites, and the
  constructors do the algebra.

  \subsection{Multiplicative Factoring}
  \label{sec:factoring}

  Canonicalisation normalises the additive structure of an expression, but it
  leaves products opaque. The find-rules of Component~I
  (Sections~\ref{sec:simple-rule} and \ref{sec:general-substitution}) look for a
  random in \emph{additive} position, as an operand of a sum, because that is the
  shape Lemma~\ref{lem:substitution} rewrites. A random that occurs only inside
  products is invisible to them, even when the expression could in fact be
  discharged. Multiplicative factoring is the pass that reshapes such an
  expression. It pulls a shared factor out of a sum of products so that a trapped
  random surfaces additively, where a substitution can reach it. It is the second
  tier of the engine and is interleaved into the third
  (Chapter~\ref{chap:design}).

    \subsubsection{The Factoring Transformation}
    \label{sec:factoring-transformation}

    Factoring applies the distributive law right to left, over a sum whose
    operands are products. If a node $f$ appears as an operand in two or more of
    those products, we pull it out,
    \[
      (f \odot a) \;\oplus\; (f \odot b) \;\oplus\; \mathit{rest}
      \;=\;
      \big(f \odot (a \oplus b)\big) \;\oplus\; \mathit{rest},
    \]
    where $\mathit{rest}$ are the summands not carrying $f$, and $a, b$ are what
    remains of the two products once $f$ is removed. Over $\mathbb{F}_2$ this is an
    identity, so the wire's value is untouched and only its form changes.

    What we pull out is chosen greedily. Over the operands of a sum we count, for
    each candidate factor, in how many of the product summands it appears, and take
    the one with the highest count, provided it appears at least twice, since a
    factor sitting in a single product buys nothing. Ties are broken by identifier,
    for determinism. With the factor $f$ fixed, we partition the summands into
    those carrying $f$ and those not, strip $f$ from each of the former, sum the
    remainders into an inner sum, multiply that by $f$, and add the untouched
    summands back, every construction going through the canonicalising
    constructors of Section~\ref{sec:canonicalization}. Factoring is then applied
    recursively, bottom up over the DAG and repeated on its own result, until no
    sum has a factor of multiplicity two or more.

    For a concrete instance we take a probe from the first-order masked quadratic
    bijection $Q^{4}_{12}$, one of the gadgets we verify. Its inputs are shared as
    $a_1 = a \oplus r_0$, $b_1 = b \oplus r_1$, $c_1 = c \oplus r_2$ together with
    the mask shares $b_2 = r_1$ and $c_2 = r_2$, where $a, b, c$ are secret and
    $r_0, r_1, r_2$ are fresh randoms. One of the gadget's recombination wires
    gathers five product and share terms, and after canonicalisation its value is
    \[
      w \;=\; (a \oplus r_0)\odot(b \oplus r_1) \;\oplus\; (a \oplus r_0)\odot(c \oplus r_2)
          \;\oplus\; (c \oplus r_2) \;\oplus\; (a \oplus r_0)\odot r_1
          \;\oplus\; (a \oplus r_0)\odot r_2 .
    \]

    As it stands the simple rule cannot finish. It clears $r_0$, whose only
    occurrence is the additive one in $a \oplus r_0$, but $b$ and $c$ then remain,
    and neither $r_1$ nor $r_2$ is a single-occurrence random. Each shows up both
    additively, in a share, and again inside the products, so the uniqueness the
    simple rule needs is absent and the general rule's side condition fails as well.
    Factoring is what breaks the deadlock. The factor $a \oplus r_0$ is shared by
    four of the five summands, and pulling it out drops their cofactors into a
    single inner sum,
    \[
      w \;=\; (a \oplus r_0) \odot \big[(b \oplus r_1) \oplus (c \oplus r_2) \oplus r_1 \oplus r_2\big]
          \;\oplus\; (c \oplus r_2)
        \;=\; (a \oplus r_0)\odot(b \oplus c) \;\oplus\; c \;\oplus\; r_2 ,
    \]
    where inside the bracket $r_1 \oplus r_1$ and $r_2 \oplus r_2$ cancel by the
    additive law of Section~\ref{sec:canonicalization}. The surplus copies of
    $r_2$ go with them, and what is left carries $r_2$ additively and exactly once.
    Now the single-occurrence pattern of Section~\ref{sec:simple-rule} applies. The
    engine relabels the whole expression as a fresh random and reads $w$ as
    secret-free.

    \subsubsection{Enlarging the Class of Verifiable Circuits}
    \label{sec:enlarging-verifiable}

    Factoring is sound by construction, being an $\mathbb{F}_2$ identity, so it
    never lets an insecure probe pass. What it changes is completeness, and it
    strictly enlarges the class of probes the engine can certify. Every probe the
    engine discharges without factoring it still discharges, since
    Algorithm~\ref{alg:rewrite-engine} tries the unfactored form first and factors
    only on failure. And some probes, like the one above, are discharged only once
    factoring has exposed a random no additive rule could otherwise reach. The
    verifiable class with factoring is therefore a strict superset of the class
    without it.

    Two cautions temper the picture, both settled already in
    Chapter~\ref{chap:design}. First, factoring is not monotone in the cheap
    direction. It can just as well destroy a single-occurrence opportunity that the
    unfactored form exposed, which is exactly why the engine tries the simple rule
    unfactored before it factors, rather than factoring once up front
    (Section~\ref{sec:simple-rule}). It has to be interleaved with substitution,
    not run as a preprocessing step. Second, factoring is not a replayable
    operation. Re-applying it to an unrelated candidate wire does not carry over
    the blinding the way re-applying a substitution does, and this is what forces
    Component~III to read a derivation as a coupling with an exact endpoint test
    rather than as a script to rerun (Section~\ref{sec:coupling-extension}). The
    gain in verifiable circuits is real, but it is bought with the extra machinery
    those sections build around it.

\section{Implementing the Rewrite Engine}
\label{sec:rewrite-engine-impl}

Algorithm~\ref{alg:rewrite-engine} presents Component~I as a single tiered
procedure, but the tool realises its rule set twice, and the reason for the two
is worth an explanation. One is a cheap, verdict-only path that answers
secret-independence without ever rewriting the DAG, and the other is a complete
path that rewrites for real and hands back the coupling. The tiering of
\textproc{CheckProbe} threads them together. It tries the cheap path first on the
unfactored probe, then once more after factoring, and only then falls through to
the complete loop. The two paths
answer different callers. Component~II's large-set pruning
(Section~\ref{sec:large-set-pruning}) throws the big union at the engine and wants
only a yes or no, as fast as possible, whereas a certified representative must
come back with the coupling that Component~III will use
(Section~\ref{sec:coupling-extension}). The next two subsections take the paths
in turn.

  \subsection{Reference-Counted Single-Occurrence Discharge}
  \label{sec:refcount-impl}

  The cheap path applies the simple rule (Section~\ref{sec:simple-rule}) to a
  fixpoint without ever building a new node. It relies on the observation that the
  simple rule needs to know very little about a random. It needs how many times
  the random occurs and whether that single occurrence is additive, and it needs
  the parent to splice away. So instead of rewriting the expression the path keeps
  that bookkeeping and updates it in place.

  A single depth-first pass over the probe's sub-DAG records, for every node, how
  many parents it has and how many of them are sums, together with its
  multiplicative depth. A random is a simple-rule candidate exactly when it has
  one parent and that parent is a sum, and it is not itself a probe root; the
  candidates are held in a todo list ordered by increasing multiplicative depth, so
  the shallowest gets rewritten first. Firing the rule on a random is then a local edit to
  the counts and not to the DAG. We mark its additive parent as freshly random,
  drop the random, and decrement the parent counts of the parent's other children.
  A sibling whose total count thereby falls to a single additive occurrence
  cascades back into the worklist as a new candidate, and it is this cascade that
  carries the fixpoint forward. The probe is secure once no secret is reachable
  from any root.

  Keeping observed wires out of the todo list is an important soundness condition.
  A random that is itself a probe root must never be relabelled,
  since its value is exactly what the adversary sees, and marking its parent fresh
  while conditioning on it would be mathematically unsound. The
  cascade re-inserts a candidate only after the same root check.

  What this path cannot do is produce a coupling. Its cascade steps fire on
  already-relabelled parents rather than on tape variables, so they are not the
  leaf relabels that Section~\ref{sec:coupling-view} assembles a coupling from, and
  replaying them against another wire would diverge from the certified derivation.
  This costs nothing where the path is used. The large-set prune only needs the
  verdict, and a probe that must yield a coupling is sent to the complete loop
  instead anyways.

  \subsection{Recomputing the Find-Rules per Round}
  \label{sec:findrule-impl}

  The complete loop (\textproc{GeneralRewrite}, Section~\ref{sec:general-substitution})
  substitutes for real, as in it alters the DAG. It picks a random and an additive occurrence
  $x = e \oplus r$ and rewrites the probe by $r \mapsto e \oplus r$ throughout,
  re-interning the result through the canonicalising constructors of
  Section~\ref{sec:canonicalization}, which is what makes the chosen occurrence
  collapse to a bare $r$ while every other absorbs $e$. Because the substitution
  genuinely reshapes the expression, the incremental counts of
  Section~\ref{sec:refcount-impl} cannot be carried across it. So the loop takes
  the opposite stance and recomputes. Each round runs one fresh depth-first pass
  over the current form, and that single pass does double duty. It answers whether
  any secret still survives, which is the loop's success test, and it produces the
  parent counts and depths the find-rules need. The pass is restricted to the
  randoms actually present in the probe rather than every random in the circuit.

  On that snapshot, the loop first tries \textproc{FindSimple}, the same
  single-occurrence additive pattern as before, since a general substitution can
  turn a twice-occurring random into a once-occurring one; failing that,
  \textproc{FindGeneral} looks for a not-yet-used random with an additive parent
  whose other children do not mention it, which is the side condition
  Lemma~\ref{lem:substitution} needs. That last check is a reachability question,
  does $r$ occur anywhere below the context $e$, answered by a memoised walk whose
  memo is shared across the candidate's parents so the sub-DAG is not re-explored
  from scratch. Both rules prefer the shallowest candidate. When both stall the
  loop factors once and retries (Section~\ref{sec:factoring}), and a fuel counter
  backstops the whole thing, as Section~\ref{sec:general-substitution} discussed.

  The point of paying this recomputation is the record it leaves. Every
  substitution the loop commits is a leaf relabel $(r,\, e \oplus r)$, and appended
  in order these are exactly the shears $\sigma_1, \dots, \sigma_K$ of
  Section~\ref{sec:coupling-view}. The loop hands that list back alongside its
  verdict, and it becomes the coupling Component~III replays.

\section{Evaluating the Coupling}
\label{sec:coupling-impl}

Component~III decides, for each candidate wire $u$, whether $u \circ \varphi$ is
free of secrets, where $\varphi$ is the coupling the engine just recorded
(Section~\ref{sec:coupling-extension}). Algorithm~\ref{alg:extend} gives that
decision, together with a probabilistic filter first and an exact test after. Both stages
need to evaluate a wire through the coupling, and this section describes how that
evaluation is done cheaply.

  \subsection{Bit-Sliced Evaluation Through a Coupled Environment}
  \label{sec:bitslice-impl}

  The filter compares two runs of $u \circ \varphi$, one with the secrets drawn at
  random and one with them zeroed, and rejects $u$ when they disagree
  (Section~\ref{sec:coupling-extension}). A single pair of runs proves little, so
  the filter wants many, and the representation makes many nearly as cheap as one.
  Every wire value is carried as a machine word whose $64$ bits are $64$
  independent $\mathbb{F}_2$ evaluations, one per lane. A sum is a word
  exclusive-or and a product a word and, so one traversal of the sub-DAG evaluates
  all $64$ lanes at once, and node values are cached across the traversal so a
  shared subexpression is computed a single time.

  Evaluating $u \circ \varphi$ never rebuilds $u$. Recall the coupling is a list
  of relabels $r \mapsto e \oplus r$; rather than substitute them into $u$, we
  precompute a \emph{coupled environment} that binds each random to the value of
  its net image under the whole list and then evaluate $u$ against it. Since a
  relabel's image may mention randoms that later relabels move, the environment is
  built by folding the list from right to left, so that by the time a random is
  bound, every relabel acting on its image is already computed. The identity we get
  is $\mathrm{eval}(u,\ \mathit{env}) = \mathrm{eval}(u \circ \varphi)$, so the two
  environments, secrets-as-random and secrets-as-zeros, are each built once and then
  reused for every candidate, with their evaluation caches shared across.

  % This felt like it was unnecessarily detailed, so I'll comment it out.
  % One subtlety decides whether the filter rejects anything at all. The lane values
  % handed to the leaves must look random in every bit. A plain linear generator
  % will not do, since the low bit of its state merely alternates and the next cycles
  % with period four, so the low lanes assign leaves highly structured, colliding
  % values and catch almost no dependence. We pass each generated word through a
  % bit-mixing finaliser first, which makes every output bit a function of the whole
  % state, so all $64$ lanes reject independently. Publics are evaluated as leaves
  % here, not folded to constants, since a dependence that shows only jointly with a
  % public would otherwise slip past the filter.

  \subsection{From Probabilistic Filter to Exact Verdict}
  \label{sec:anf-impl}

  Agreement across lanes is solid evidence but not a proof, so the filter is allowed to
  reject but never to accept (Lemma~\ref{lem:extend-exact}). A survivor has to be decided
  exactly. The coupling is applied to all survivors together, sharing one
  substitution memo per step rather than replaying the whole list per candidate,
  and each image is then read in two ways. First by ordinary $\mathbb{F}_2$
  cancellation over the canonical form, which is cheap and clears the common case.
  When a secret still shows there, the image is converted to its algebraic normal
  form, and the wire is admitted exactly when no secret variable occurs in its ANF.
  The ANF is the canonical multilinear form (Section~\ref{sec:field-f2}), so a
  secret is absent from it precisely when the wire does not depend on that secret;
  this is what makes the test exact rather than merely sound.

  % Again a little too detail here as well, so I'll comment for now.
  % The one hazard is cost, since an ANF can be exponentially large. The conversion
  % therefore carries a monomial budget and aborts once it is exceeded, and an
  % aborted conversion rejects the candidate. Rejection only ever costs extension
  % yield and never soundness, because an admission always rests on a completed,
  % exact form and never on the probabilistic filter or on a budget guess. As in the
  % loops above, the work is shared. The monomial expansions are memoised by node
  % across the survivors, which typically overlap heavily in their sub-DAGs, so the
  % exact stage pays for each shared subexpression once.
