\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Conclusion}
\label{chap:conclusion}
% ------------------------------------------------------------
Masking is prone to design flaws. Shares of a secret might entangle as non-linearity
necessitates it to a degree. Hence, a checker verifying that too many shares never come back
together is of value. This thesis set out to make that check exact. The question we fixed
on is $d$-probing security of masked arithmetic circuits over $\mathbb{F}_2$,
checking whether every set of at most $d$ wires is jointly independent of the secrets, and
our aim throughout was to answer it without approximation, either certifying a
circuit or handing back a concrete probe that we suspect to leak. The difficulty is that
exactness is expensive, and much of the work was in finding a route that pays for
it in an accaptable way.

We began by measuring the cost precisely. The law of a probe is a histogram of tape
counts, and we showed that computing one such count is $\#\mathrm{P}$-complete while
deciding whether two secrets leave the law unchanged is
$\mathrm{C}_{=}\mathrm{P}$-complete (Chapter~\ref{chap:couplings}). Even a single
probe therefore lands above the polynomial hierarchy, which settles that no exact
procedure built on mere counting can be efficient and points away from the distribution
and toward its structure. This is the motivation that shapes our approach.

The rest of the thesis is the constructive answer, an executable checker,
\toolname, organised around three components and one idea that ties them together.
Heavily based on Barthe et al.'s work on language-based verification
\parencite{barthe2015verified, 10.1007/978-3-030-29959-0_15}.
Component~I, the rewrite engine, decides a single probe by substituting its randoms
away one at a time until no secret remains, cheaply when each random occurs once
and by a generalized fallback otherwise, with a multiplicative factoring step that
reaches masks buried inside products (Chapter~\ref{chap:design}). Component~II
sweeps every $d$-subset without enumerating them, splitting the space recursively
and losing nothing, which we proved by Vandermonde's identity at the level
of sets. Component~III grows a single certified probe into a whole safe set of
wires, so that one discharge settles many probes at once. The idea underneath all
three is the \emph{coupling}. A successful discharge is not merely a secure verdict, but a
measure-preserving relabelling of the randomness under which the secrets visibly
cancel, and reading it that way is what lets one certificate vouch for a probe, for
its neighbour wires, and for the soundness of the traversal that calls it. The resulting
checker is a sound but incomplete tool.

We implemented \toolname\ in Lean~4 on a hash-consed circuit representation with a
bit-sliced test for the extension (Chapter~\ref{chap:implementation}), and
evaluated it on a suite of masked gadgets (Chapter~\ref{chap:evaluation}), where it
certifies the secure constructions and, on an insecure one, returns a witnessing
probe.

What we take from the work is that couplings are the right lens on probing
security. They turn a distributional statement into a constructive object one can
build, record, and replay, and they unify the three components under a single
soundness argument rather than three. As stated, the approach stands in the language-based
lineage of Barthe et al.\ \parencite{barthe2015verified,
10.1007/978-3-030-29959-0_15}, and its departures, the explicit coupling with an
exact endpoint, the factoring, the bit-sliced fast-reject test, are what this thesis
contributes to that line. Much
is left open, from larger fields and glitches to composition and to the composing of
couplings across safe sets (Chapter~\ref{chap:futurework}), and the hardness result
guarantees that no single technique will close the gap completely.
