\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Couplings and Probing Security}
\label{chap:couplings}
% ------------------------------------------------------------
\section{The Counting Problem is \NoCaseChange{\#P}-Complete} % this is at a weird place, but I want to mention it
\todo[inline]{Introduce the counting problem behind secret-independence checking and state the \#P-completeness result.}
  \subsection{From Independence to Distribution Counting}
  \todo[inline]{Show how deciding secret-independence of a probe reduces to a distribution-counting problem.}
  \subsection{\NoCaseChange{\#P}-Completeness and Its Consequences for Verification}
  \todo[inline]{Prove \#P-completeness of the counting problem and discuss what this implies for exact verification (motivating the constructive rewrite-based approach instead).}
\section{Couplings of Probabilistic Programs}
\todo[inline]{Introduce couplings as a proof technique for distributional equality between probabilistic programs.}
  \subsection{Coupling Two Runs Under Different Secrets}
  \todo[inline]{Define a coupling between the run of a probe under one secret assignment and its run under another.}
  \subsection{Coupling Functions: Randoms to Randoms} % measure-preserving bijections! on randomness
  \todo[inline]{Formalize coupling functions as measure-preserving bijections on the randomness domain.}
\section{Probing Security via Couplings}
\todo[inline]{Relate $d$-probing security to the existence of a coupling for every probe of size $d$.}
  \subsection{\NoCaseChange{$d$}-Probing Security Implies a (Family of) Coupling Functions}
  \todo[inline]{Prove that $d$-probing security is equivalent to the existence of a family of coupling functions, one per pair of secret assignments.}
  \subsection{Secret-Indistinguishability}
  \todo[inline]{Connect the coupling-existence characterization back to the standard secret-indistinguishability definition of probing security.}
\section{Substitutions as a Coupling Construction}
\todo[inline]{Explain how the randomness-substitution rewrites of Section~\ref{sec:randomness-substitution} construct an explicit coupling.}
  \subsection{The Derivation of an Explicit Coupling Witness}
  \todo[inline]{Show how a derivation (sequence of substitutions) yields an explicit coupling witness for a probe.}
  \subsection{Why a Successful Rewrite Yields a Valid Coupling}
  \todo[inline]{Prove that a successful rewrite derivation, read as a composition of shears, satisfies the coupling definition.}
