\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Design of \toolname}
\label{chap:design}
% ------------------------------------------------------------
\section{Pipeline Overview}
\label{sec:pipeline-overview}
\todo[inline]{Give a high-level overview of the verification pipeline: the three components (rewrite engine, subset-space traversal, probe-set extension) and how they fit together.}
\section{Component I: The Rewrite Engine}
\label{sec:rewrite-engine}

Given a chosen probe $\mathbf{w} = (w_1, \dots, w_k)$, $k \le d$ --- a tuple of
wires of the circuit --- the rewrite engine decides whether $\mathbf{w}$ is
secret-independent. It does so constructively: rather than reasoning about
probability distributions directly, it searches for a finite sequence of
applications of Lemma~\ref{lem:substitution}
(Section~\ref{sec:randomness-substitution}) after which $\mathbf{w}$'s image is
\emph{syntactically} free of secret leaves --- no secret leaf is reachable from
any $w_i$. Termination with a secret-free image certifies $\mathbf{w}$;
Theorem~\ref{thm:rewrite-soundness} in Section~\ref{sec:coupling-view} makes precise
why this certificate is sound.

The engine has three tiers of rewriting rules, increasing in generality and cost,
summarized in
Algorithm~\ref{alg:rewrite-engine} and mirrored in the structure of this section.
\textproc{SimpleRewrite} (Section~\ref{sec:simple-rule}) handles the common case,
every random occurring at most once, cheaply. It is tried first on the unfactored
probe and, on failure, once more after the multiplicative factoring of
Section~\ref{sec:factoring} has had a chance to expose additional single-occurrence
randoms. \textproc{GeneralRewrite} (Section~\ref{sec:general-substitution}) is the
complete fallback for the remaining case, a random occurring more than once: it
substitutes a random by its context wherever the random occurs, interleaving
factoring lazily whenever both of its own find-rules stall. Section~\ref{sec:coupling-view}
closes the section by reading the whole derivation as an explicit, replayable
bijection on the randomness assignment, the \gls{coupling} that Component III
(Section~\ref{sec:extension}) later replays against every other wire in the circuit.

\begin{algorithm}
\caption{Discharging a probe (Component I)}
\label{alg:rewrite-engine}
\begin{algorithmic}[1]
\Function{CheckProbe}{$\mathbf{w}$}
  \Require probe $\mathbf{w}$, a tuple of wires of the circuit
  \Ensure \textsc{Secure}, or \textsc{Insecure}$(\mathbf{w})$
  \State $(\mathit{ok}, T) \gets \Call{SimpleRewrite}{\mathbf{w}}$ \Comment{Tier 1: unfactored}
  \If{$\mathit{ok}$} \Return \textsc{Secure} \EndIf
  \State $(\mathit{ok}, T) \gets \Call{SimpleRewrite}{\mathrm{Factor}(\mathbf{w})}$ \Comment{Tier 2: factor (\S\ref{sec:factoring}), retry}
  \If{$\mathit{ok}$} \Return \textsc{Secure} \EndIf
  \State $(\mathit{ok}, T) \gets \Call{GeneralRewrite}{\mathbf{w}, \emptyset, \mathit{fuel}, \mathit{false}, \langle\rangle}$ \Comment{Tier 3: complete fallback}
  \State \Return $\mathit{ok}$ ? \textsc{Secure} : \textsc{Insecure}($\mathbf{w}$)
\EndFunction
\Statex
\Function{SimpleRewrite}{$\mathit{roots}$}
  \State $\mathit{todo} \gets \{\, r \in \mathrm{vars}(\mathit{roots}) : r \text{ occurs once, additively}, r \notin \mathit{roots} \,\}$, shallowest first
  \State $T \gets \langle\rangle$
  \While{$\mathit{todo} \neq \langle\rangle$ \textbf{and not} \Call{SecretFree}{$\mathit{roots}$}}
    \State $r \gets$ pop shallowest element of $\mathit{todo}$
    \State \ $x \gets$ unique additive parent of $r$ \Comment{$x = e \oplus r$}
    \State mark $x$ as a fresh random leaf
    \State \ $T \gets T.\mathrm{push}(r, x)$ \Comment{$t = e \oplus r = x$, already in hand}
    \State drop $r$ from the tuple
    \State \emph{(a random newly occurring once, additively, and $\notin \mathit{roots}$, joins $\mathit{todo}$)}
  \EndWhile
  \State \Return $(\Call{SecretFree}{\mathit{roots}}, T)$ \Comment{same shears, now recorded --- \S\ref{sec:coupling-view}}
\EndFunction
\Statex
\Function{GeneralRewrite}{$\mathit{tuple}, \mathit{used}, \mathit{fuel}, \mathit{factored}, T$}
  \If{\Call{SecretFree}{$\mathit{tuple}$}} \Return $(\mathit{true}, T)$ \EndIf
  \If{$\mathit{fuel} = 0$} \Return $(\mathit{false}, T)$ \EndIf
  \State $r \gets \Call{FindSimple}{\mathit{tuple}}$ \Comment{occurs once, additively, not a root}
  \If{$r = \bot$} \ $r \gets \Call{FindGeneral}{\mathit{tuple}, \mathit{used}}$ \Comment{occurs $\geq 2\times$; some additive parent free of $r$} \EndIf
  \If{$r = \bot$}
    \If{$\mathit{factored}$} \Return $(\mathit{false}, T)$ \EndIf
    \State \Return $\Call{GeneralRewrite}{\mathrm{Factor}(\mathit{tuple}), \mathit{used}, \mathit{fuel}, \mathit{true}, T}$ \Comment{\S\ref{sec:factoring-cost}, lazy retry}
  \EndIf
  \State $x \gets$ chosen additive parent of $r$; \ $e \gets \bigoplus(\mathrm{children}(x) \setminus \{r\})$ \Comment{side cond.\ $r \notin \mathrm{vars}(e)$}
  \State $t \gets e \oplus r$; \ $\mathit{tuple}' \gets \mathit{tuple}[r \mapsto t]$ \Comment{substitute everywhere --- Lemma~\ref{lem:substitution}}
  \State \Return $\Call{GeneralRewrite}{\mathit{tuple}', \mathit{used} \cup \{r\}, \mathit{fuel}-1, \mathit{false}, T.\mathrm{push}(r,t)}$
\EndFunction
\end{algorithmic}
\end{algorithm}

  \subsection{The Simple Rewrite Rule}
  \label{sec:simple-rule}

  The cheapest tier discharges the pattern
  Corollary~\ref{cor:single-occurrence-discharge} addresses directly: a random that
  occurs exactly once below the probe, additively --- as an operand of an XOR gate.
  Such a random is a \emph{simple-rule candidate}, provided it is not itself a probe
  root. Writing $x = e \oplus r$ for its unique parent, the corollary lets us relabel
  $x$ as a fresh random and drop $r$ without ever looking at the context $e$: the
  pattern alone certifies that $x$, jointly with everything independent of $r$, is
  distributed as fresh, independent randomness. The tier applies this rewrite
  repeatedly, each discharge possibly leaving a further random occurring once, until
  either no secret leaf is reachable from the probe --- success --- or no candidate
  remains --- failure.

  The exclusion of probe roots is a soundness condition, not an optimisation. A
  root's value is what the adversary observes: it is fixed, not fresh, and relabelling
  it would claim independence for something the tuple actually exposes. The tuple
  $(r,\ s \oplus r,\ r \oplus r_3)$ makes the danger concrete. Its first entry is the
  bare leaf $r$, and the first two entries XOR to the secret $s$, so this tuple leaks
  $s$ and must be rejected. Yet once $r_3$ is discharged --- it occurs once,
  additively, in $r \oplus r_3$ --- the random $r$ is left occurring only once, in
  $s \oplus r$. Were roots not guarded, $r$ would now match the candidate pattern and
  be relabelled away, blinding the very value the adversary reads and reporting a
  false \textsc{Secure}. Guarding roots throughout the cascade is what rules this out.

  Two properties of the tier matter downstream. First, it fires the shallowest,
  closest-to-a-root candidates first, matching the increasing-depth discharge order
  of \texttt{maskVerif} \parencite{10.1007/978-3-030-29959-0_15}. Second, a random is
  discharged at most once --- relabelled, it is gone --- so the tier performs at most
  $|\mathcal{R}|$ rewrites and always terminates. Its realisation as an incremental
  traversal that avoids recomputing the candidate set after each discharge is deferred
  to Section~\ref{sec:rewrite-engine-impl}.

% refer to the soundness proof
  This tier is sound, since each step is a direct instance of
  Corollary~\ref{cor:single-occurrence-discharge}, but incomplete: it has nothing to
  offer a random occurring twice, such as a mask reused across two cross-terms of a
  masked multiplication. Section~\ref{sec:general-substitution}'s fallback handles
  that case.

  \subsection{Fallback: General Substitution}
  \label{sec:general-substitution}

  \textproc{GeneralRewrite} handles randoms that occur more than once by performing
  the substitution of Lemma~\ref{lem:substitution} literally, rather than only its
  single-occurrence case (Corollary~\ref{cor:single-occurrence-discharge}). It picks
  a random $r$ and an additive occurrence $x_1 = e \oplus r$ with
  $r \notin \mathrm{vars}(e)$, then rewrites every root by substituting
  $r \mapsto e \oplus r$ throughout. Cancellation in $\mathbb{F}_2$ collapses the
  chosen occurrence $x_1$ to the bare leaf $r$, exactly as in
  Corollary~\ref{cor:single-occurrence-discharge}, while every other occurrence of
  $r$ absorbs $e$ instead of vanishing, cancelling further if it happens to share
  structure with $e$. The loop succeeds once no secret leaf remains reachable from
  the rewritten roots.

  Each round re-examines the current tuple for an applicable substitution.
  \textproc{FindSimple} looks first for the same single-occurrence, non-root pattern
  as Section~\ref{sec:simple-rule}, since a general substitution can turn a
  multi-occurrence random into a single-occurrence one; it prefers the shallowest
  candidate. Failing that, \textproc{FindGeneral} looks for any random, not yet used
  by a previous general step, occurring at least twice with some additive parent
  $x_1$ whose other children do not themselves depend on it --- the side condition
  $r \notin \mathrm{vars}(e)$ that Lemma~\ref{lem:substitution} requires. It prefers
  the deepest candidate, the opposite bias from \textproc{FindSimple}; no deeper
  principle motivates this choice beyond the ordering itself. A context $e$ is always
  read off an XOR gate's other children: anything else would mean $e$ contains $r$,
  making $r \mapsto e \oplus r$ fail to be a bijection, so such a context is rejected
  outright rather than admitted as an unsound relabelling.

  When both find-rules stall on the tuple's current form, the engine factors
  (Section~\ref{sec:factoring}) once and retries, rather than declaring failure
  immediately. Factoring can move a random from a purely multiplicative position,
  where neither find-rule can see it, into an additive one, but on its own it
  eliminates no secret, so it must be interleaved with substitution rather than run
  once up front or once at the end. Factoring is symmetrically not tried before the
  simple rule either: Algorithm~\ref{alg:rewrite-engine} always tries
  \textproc{SimpleRewrite} unfactored first, because factoring can destroy a
  single-occurrence opportunity the unfactored form exposes, so trying the cheap,
  unfactored tier first is not just faster on the common case but strictly more
  complete.

  \textproc{FindGeneral} is needed precisely when a random occurs additively in two
  places that do not telescope into one another under the simple rule. Suppose $r$
  occurs as an operand of $x_1 = e_1 \oplus r$ in one root and of $x_2 = e_2 \oplus r$
  in another, with $r \notin \mathrm{vars}(e_1) \cup \mathrm{vars}(e_2)$.
  Substituting through $x_1$, i.e.\ $r \mapsto e_1 \oplus r$, turns $x_1$ into
  $e_1 \oplus (e_1 \oplus r) = r$ by cancellation, as in
  Corollary~\ref{cor:single-occurrence-discharge}, while the same substitution turns
  $x_2$ into $e_2 \oplus e_1 \oplus r$. If $e_1 \oplus e_2$ is itself secret-free,
  both roots are secret-free after this one step; otherwise the loop continues,
  substituting further or factoring to expose a new opportunity. This is the shape of
  the case a masked multiplication typically produces, where a single random masks
  two different cross-terms: neither term's random is disposable on its own, but the
  substitution above can still cancel the secret-dependent part they share. Whether a
  given instance is discharged this way, and in how many rounds, depends entirely on
  how much of $e_1 \oplus e_2$ the rest of the derivation can still cancel; the loop
  offers no guarantee beyond trying. Even granting unlimited fuel, it only composes
  single-coordinate shears $\widehat{\sigma}$, so it can only certify
  secret-independence witnessed by a chain of substitutions, possibly interleaved
  with factoring. Completeness is proved only in the linear/affine case
  \textcolor{orange}{(nerede proof?)}; whether a genuinely irreducible counterexample
  exists beyond it is not settled by this thesis.

  Termination is only argued informally, not proved. Each general substitution
  retires its random into $\mathit{used}$ and is never repeated for it, bounding the
  number of general steps by $|\mathcal{R}|$; between general steps, simple
  substitutions strictly reduce the candidate pool; factoring is retried at most once
  per stall before the engine gives up on the current form. Interleaving factoring
  with two mutually triggering find-rules resists a clean potential-function
  argument, so a generous fuel counter backstops the loop instead. Exhausting fuel is
  treated as failure, which is sound (Theorem~\ref{thm:rewrite-soundness} is never
  invoked on a fuel-out) but not necessarily tight: a fuel-out reports the probe as
  though it were genuinely insecure, without exhibiting an actual distinguishing
  observation.

  \subsection{Relation to the Coupling-Construction View}
  \label{sec:coupling-view}

  A discharge does more than answer \textsc{Secure}/\textsc{Insecure}. Read the
  right way, it hands back the object Chapter~\ref{chap:couplings} asks for: a
  coupling for the probe. We state that object declaratively first, then show how
  the engine produces one.

  Let the secrets range now, rather than staying fixed as in
  Section~\ref{sec:randomness-substitution}: write
  $\mathbf{w}(\rho; \mathbf{a})$ for the value of the probe under randomness
  assignment $\rho \in \{0,1\}^{\mathcal{R}}$ and secret assignment
  $\mathbf{a} \in \{0,1\}^{\mathcal{S}}$, the public inputs held fixed.

  \begin{definition}[Coupling of a probe]
  \label{def:probe-coupling}
  A \emph{coupling} of a probe $\mathbf{w}$ is a family
  $\{\,\Phi_{\mathbf{a},\mathbf{b}}\,\}_{\mathbf{a},\mathbf{b} \in \{0,1\}^{\mathcal{S}}}$
  of measure-preserving bijections of $\{0,1\}^{\mathcal{R}}$, one per ordered
  pair of secret assignments, satisfying
  \[
    \mathbf{w}(\rho; \mathbf{a}) \;=\; \mathbf{w}\big(\Phi_{\mathbf{a},\mathbf{b}}(\rho);\, \mathbf{b}\big)
    \qquad \text{for every } \rho \in \{0,1\}^{\mathcal{R}} .
  \]
  \end{definition}

  Definition~\ref{def:probe-coupling} says precisely where to send the randomness
  so that a run under $\mathbf{a}$ is reproduced, observation for observation, by
  a run under $\mathbf{b}$: each $\Phi_{\mathbf{a},\mathbf{b}}$ dictates the
  relabelling of the random tape that absorbs the change of secret. Since it is
  measure-preserving, pushing the uniform tape through it turns this pointwise
  identity into an equality of laws, so the distribution of
  $\mathbf{w}(\cdot;\mathbf{a})$ matches that of $\mathbf{w}(\cdot;\mathbf{b})$. A
  coupling existing for \emph{every} pair $(\mathbf{a},\mathbf{b})$ is thus the
  same statement as $\mathbf{w}$ being secret-independent --- which is what makes
  it the right certificate. The family carries the structure one expects of such
  a certificate: one may take $\Phi_{\mathbf{a},\mathbf{a}} = \mathrm{id}$,
  $\Phi_{\mathbf{b},\mathbf{a}} = \Phi_{\mathbf{a},\mathbf{b}}^{-1}$, and
  $\Phi_{\mathbf{b},\mathbf{c}} \circ \Phi_{\mathbf{a},\mathbf{b}} =
  \Phi_{\mathbf{a},\mathbf{c}}$, so a single anchor
  $\Phi_{\mathbf{a}_0,\,\cdot}$ already generates all of it. It is these
  properties we care about, not any one procedure that realises them.

  What the engine records is one such procedure. Discharging a probe amounts to
  finding a \emph{derivation}: a finite sequence of rewrites
  \[
    \sigma_1 = (r_1 \mapsto e_1 \oplus r_1), \ \dots, \ \sigma_K = (r_K \mapsto e_K \oplus r_K),
  \]
  each valid --- $r_i \notin \mathrm{vars}(e_i)$ --- against the tuple's state
  just before step $i$, after which no secret leaf remains reachable. A rewrite
  is a rewrite: whatever rule proposed it, and whether the random it retires
  occurred once or many times, each $\sigma_i$ is a substitution in the sense of
  Definition~\ref{def:substitution}, hence a single measure-preserving shear
  $\widehat{\sigma_i}$ (Lemma~\ref{lem:substitution}). The coupling is assembled
  from these shears and from nothing else, so the construction below never needs
  to know how any individual rewrite was found.

  \begin{theorem}[Soundness of discharge]
  \label{thm:rewrite-soundness}
  If \textup{\textproc{CheckProbe}}$(\mathbf{w})$
  (Algorithm~\ref{alg:rewrite-engine}) returns \textup{\textsc{Secure}}, then the
  joint distribution of $\mathbf{w}(\rho; \mathbf{a})$, for $\rho$ sampled
  uniformly at random, does not depend on the secret assignment $\mathbf{a}$.
  Moreover the recorded substitutions realise a coupling of $\mathbf{w}$ in the
  sense of Definition~\ref{def:probe-coupling}.
  \end{theorem}

  \begin{proof}[Proof sketch]
  Fix a secret assignment $\mathbf{a}$. Each recorded step contributes the tape
  shear $\widehat{\sigma_i}$, whose defining context $e_i$ may itself read the
  secrets, so write $\widehat{\sigma_i}^{\,\mathbf{a}}$ for its action under
  $\mathbf{a}$ and set
  \[
    \varphi_{\mathbf{a}} \;=\; \widehat{\sigma_1}^{\,\mathbf{a}} \circ
    \widehat{\sigma_2}^{\,\mathbf{a}} \circ \cdots \circ \widehat{\sigma_K}^{\,\mathbf{a}},
  \]
  a composition of measure-preserving bijections, hence itself one. Writing
  $\mathbf{w}_K = \mathbf{w}\sigma_1 \cdots \sigma_K$ for the final tuple and
  chaining Lemma~\ref{lem:substitution} through the $K$ steps gives
  $\mathbf{w}\big(\varphi_{\mathbf{a}}(\rho); \mathbf{a}\big) = \mathbf{w}_K(\rho)$
  for every $\rho$. A \textsc{Secure} verdict means $\mathbf{w}_K$ is
  syntactically secret-free --- no secret leaf is reachable from it --- so
  $\mathbf{w}_K(\rho)$ does not depend on $\mathbf{a}$. The right-hand side is
  therefore common to all secret assignments, and for any two, $\mathbf{a}$ and
  $\mathbf{b}$,
  \[
    \mathbf{w}\big(\varphi_{\mathbf{a}}(\rho); \mathbf{a}\big) \;=\; \mathbf{w}_K(\rho)
    \;=\; \mathbf{w}\big(\varphi_{\mathbf{b}}(\rho); \mathbf{b}\big) .
  \]
  Substituting $\rho \mapsto \varphi_{\mathbf{a}}^{-1}(\rho)$ and setting
  $\Phi_{\mathbf{a},\mathbf{b}} = \varphi_{\mathbf{b}} \circ
  \varphi_{\mathbf{a}}^{-1}$ gives
  $\mathbf{w}(\rho; \mathbf{a}) = \mathbf{w}\big(\Phi_{\mathbf{a},\mathbf{b}}(\rho); \mathbf{b}\big)$
  for all $\rho$; the family $\{\Phi_{\mathbf{a},\mathbf{b}}\}$ is the promised
  coupling, and measure-preservation of each member yields the distributional
  claim. The chaining uses only that each $\sigma_i$ is a substitution
  (Definition~\ref{def:substitution}); a single-occurrence rewrite, where $e_i$
  is read off one $\oplus$-parent as in
  Corollary~\ref{cor:single-occurrence-discharge}, is folded in exactly like any
  other, so the argument is indifferent to how the derivation was produced.
  \end{proof}

  The composite $\varphi_{\mathbf{a}}$ is displayed above as a product of shears
  only for exhibition; it is never used. Component III (Section~\ref{sec:coupling-extension})
  uses a single member of the family as a \emph{blinding certificate}: it takes any
  other wire $u$ in the circuit and admits it into the certified \emph{safe set}
  when $u \circ \varphi_{\mathbf{a}}$ is secret-free --- when the same relabelling
  that blinds the probe blinds $u$ too.
  Evaluating $\varphi_{\mathbf{a}}$ on a sampled tape is a back-substitution
  through the recorded steps in reverse order $i = K, \dots, 1$, since the net
  image of $r_i$ is $t_i = e_i \oplus r_i$ evaluated against the assignment
  already updated by every later step: the derivation performs forward
  elimination, and reading the net map off is back-substitution through the same
  steps in reverse.

  Keeping this record costs nothing. Each rewrite already names its target
  $t_i = e_i \oplus r_i$ --- the very wire it substitutes for $r_i$ --- so the
  derivation $\sigma_1, \dots, \sigma_K$ can be accumulated as the rewrites happen,
  and the coupling is read off that one list. Component III replays these shears
  directly against every other wire in the circuit; no probe and no rewrite is
  special-cased, and each step that helped certify the probe is a step that can help
  blind its neighbours.

\section{Component II: Traversing the Subset Space}
\label{sec:subset-space}

$d$-probing security asks that \emph{every} set of at most $d$ wires be
secret-independent (Section~\ref{sec:d-probing}). Read literally, that is
$\sum_{i=0}^{d} \binom{N}{i}$ separate questions for a circuit of $N$ wires.
Component II is the
bookkeeping that keeps the question exact while never enumerating a subset
redundantly.

Two elementary observations are helpful to understand the traversal.
First, we should see that secret-independence passes to
subsets, specifically a subset of a secret-independent set is a marginal of a
secret-independent law, hence itself is secret-independent. Certifying a set
of wires therefore certifies all of its subsets in one stroke. 
Second, it is enough to look at probes of size
\emph{exactly} $d$: a smaller probe is a marginal of some size-$d$ probe
(whenever $N \ge d$, which always holds here), so it is covered once the size-$d$
probes are. The task is thus to certify every $d$-element subset of the wire set,
in batches as large as possible rather than one at a time.

  \subsection{The Representation of the Subset Space}
  \label{sec:represent-subsets}

  The traversal never holds an explicit list of subsets; it holds a compact
  description of a whole family of them at once. A \emph{factor} is a pair
  $(c, W)$ of a count $c \in \mathbb{N}$ and a wire set $W$, standing for the
  family $\binom{W}{c}$ of all $c$-element subsets of $W$. A \emph{worklist} is a
  finite sequence of factors
  \[
    wl = \big\langle (c_1, W_1), \dots, (c_m, W_m) \big\rangle,
    \qquad W_1, \dots, W_m \text{ pairwise disjoint,}
  \]
  denoting the probes obtained by drawing $c_j$ wires from each $W_j$ and taking
  the union:
  \[
    \llbracket wl \rrbracket \;=\;
      \Big\{\, S_1 \cup \dots \cup S_m \;:\; S_j \in \tbinom{W_j}{c_j} \,\Big\},
    \qquad
    \big|\llbracket wl \rrbracket\big| \;=\; \prod_{j=1}^{m} \binom{|W_j|}{c_j}.
  \]
  Disjointness of the $W_j$ makes that union a disjoint one, so every probe has
  size $\sum_j c_j$ and none is described twice. Two vacaous worklists bound
  the recursion: if some factor asks for more than it holds, $|W_j| < c_j$, then
  $\binom{W_j}{c_j} = \emptyset$ and $\llbracket wl \rrbracket$ is empty; if every
  $c_j = 0$, the sole member is the empty probe. Either way there is nothing to
  certify. The verification task itself is the single-factor worklist
  $\big\langle (d, \mathcal{W}) \big\rangle$, whose denotation is
  $\binom{\mathcal{W}}{d}$, i.e. all $d$-subsets of the circuit's wires
  $\mathcal{W}$.

  Algorithm~\ref{alg:subset-space} sweeps this representation with three moves:
  discharge the union of the whole worklist and, on success, prune the branch;
  otherwise certify a representative probe and extend it to a safe set; then split
  each factor about that safe set and recurse on the pieces. The next two
  subsections treat the moves prune-and-extend and split, and presents a proof
  that the split loses nothing.

\begin{algorithm}
\caption{Traversing the subset space (Component II)}
\label{alg:subset-space}
\begin{algorithmic}[1]
\Function{CheckAll}{$wl = \langle (c_1, W_1), \dots, (c_m, W_m) \rangle$}
  \If{$\llbracket wl \rrbracket = \emptyset$} \Return \textsc{Secure} \Comment{some $|W_j| < c_j$, or every $c_j = 0$} \EndIf
  \State $U \gets W_1 \cup \dots \cup W_m$
  \If{\Call{CheckProbe}{$U$} $=$ \textsc{Secure}} \Return \textsc{Secure} \Comment{large-set pruning (\S\ref{sec:large-set-pruning})} \EndIf
  \State $\mathbf{w} \gets$ a representative of $\llbracket wl \rrbracket$ \Comment{$c_j$ wires from each $W_j$}
  \State $(\mathit{sec}, T) \gets \Call{CheckProbe}{\mathbf{w}}$ \Comment{Component I (\S\ref{sec:rewrite-engine})}
  \If{\textbf{not} $\mathit{sec}$} \Return \textsc{Insecure}($\mathbf{w}$) \EndIf
  \State $\mathsf{safe} \gets \Call{Extend}{\mathbf{w}, T}$ \Comment{Component III (\S\ref{sec:extension}); \ $\mathbf{w} \subseteq \mathsf{safe}$}
  \ForAll{$(i_1, \dots, i_m) \in \textstyle\prod_{j} \{0, \dots, c_j\}$ \textbf{with} $(i_j)_j \neq (c_j)_j$}
    \State $wl' \gets \big\langle\, (i_j,\ W_j \cap \mathsf{safe}),\ \ (c_j - i_j,\ W_j \setminus \mathsf{safe}) \,\big\rangle_{j=1}^{m}$ \Comment{drop zero-count parts}
    \State $r \gets \Call{CheckAll}{wl'}$
    \If{$r \neq \textsc{Secure}$} \Return $r$ \EndIf
  \EndFor
  \State \Return \textsc{Secure}
\EndFunction
\end{algorithmic}
\end{algorithm}

  \subsection{Large-Set Pruning}
  \label{sec:large-set-pruning}

  Before splitting anything, the traversal tries to discharge the whole batch at
  once. Let $U = W_1 \cup \dots \cup W_m$ be every wire the worklist can reach,
  and hand $U$ --- as a single tuple --- to the rewrite engine of Component I. If
  $U$ is certified secret-independent, so is every subset of $U$, and in
  particular every probe in $\llbracket wl \rrbracket$; the branch is secure and
  the traversal returns without descending. On some secure circuits this one test
  does almost all of the work --- most branches never split, because their union
  already discharges.

  However this discharge may fail, and it is worth seeing why the algorithm cannot stop here on
  failure. The union $U$ is typically far larger than $d$, and a large tuple can
  be secret-\emph{dependent} while every one of its size-$d$ sub-tuples is
  secret-independent --- so failing to discharge $U$ says nothing about the
  individual probes. When the union fails, we descend. We pick a
  \emph{representative} $\mathbf{w} \in \llbracket wl \rrbracket$ --- one probe of
  the target size, $c_j$ wires from each factor --- certify it with Component I,
  and, if it holds, grow it with Component III (Section~\ref{sec:extension}) into a
  \emph{safe set}
  \[
    \mathsf{safe} \;\supseteq\; \mathbf{w}, \qquad
    \text{every subset of which is secret-independent.}
  \]
  This will be further explained in Component III. Put briefly,
  one measure-preserving coupling
  blinds every wire of $\mathsf{safe}$ simultaneously, so any sub-tuple of
  $\mathsf{safe}$, of any size, is secret-independent. Because the
  representative draws at least one wire from every factor, $\mathsf{safe}$ meets
  every $W_j$. This single fact is what makes the split below both exhaustive and
  strictly progressing.

  \subsection{Completeness via Vandermonde's Identity}
  \label{sec:vandermonde}

  With a safe set in hand the split is forced by a counting identity. Fix
  $\mathsf{safe}$ and cut each factor's ground set along it,
  \[
    W_j = A_j \uplus B_j, \qquad A_j = W_j \cap \mathsf{safe}, \quad B_j = W_j \setminus \mathsf{safe}.
  \]

  \begin{lemma}[Set Vandermonde]
  \label{lem:set-vandermonde}
  For disjoint wire sets $A, B$ and any $c \ge 0$,
  \[
    \binom{A \uplus B}{c} \;=\; \biguplus_{i=0}^{c}
      \Big\{\, S_A \cup S_B \;:\; S_A \in \tbinom{A}{i},\ S_B \in \tbinom{B}{\,c-i} \,\Big\},
  \]
  a \emph{disjoint} union: each $c$-subset $S$ lies in exactly the block
  $i = |S \cap A|$. Counting both sides recovers Vandermonde's identity,
  $\binom{|A|+|B|}{c} = \sum_{i=0}^{c} \binom{|A|}{i} \binom{|B|}{\,c-i}$.
  \end{lemma}

  The lemma is immediate, a $c$-subset is determined by its safe part and its
  unsafe part, chosen independently. Stating it set-wise, and not merely as the
  numeric identity, is the point: it says not only \emph{how many} probes each
  block holds but exactly \emph{which}. Applying it in every factor and
  multiplying out,
  \[
    \llbracket wl \rrbracket \;=\; \prod_{j=1}^{m} \binom{W_j}{c_j}
    \;=\; \biguplus_{(i_1, \dots, i_m)} \ \llbracket wl_{(i_j)} \rrbracket,
    \qquad
    wl_{(i_j)} = \big\langle (i_j, A_j),\ (c_j - i_j, B_j) \big\rangle_{j=1}^{m},
  \]
  one \emph{cell} for each distribution $(i_1, \dots, i_m) \in \prod_j \{0, \dots, c_j\}$
  recording, factor by factor, how many of a probe's wires come from the safe
  side. The cells are disjoint and exhaust $\llbracket wl \rrbracket$, since every
  probe declares in each factor exactly how much of it is safe and each cell
  is again the denotation of a worklist, on which the traversal recurses directly.

  One cell needs no recursion. The \emph{all-safe} cell $(i_j)_j = (c_j)_j$
  collects the probes drawn entirely from the safe sides $A_j \subseteq
  \mathsf{safe}$ --- the probes lying inside $\mathsf{safe}$ --- and every one of
  those is secret-independent by the guarantee of Component III. The algorithm
  skips it and recurses on the rest. Skipping that particular cell means, with each
  recursing we are strictly decreasing the amount of probe sets we have to consider,
  which gives us our termination guarantee.

  \begin{theorem}[Soundness of the traversal]
  \label{thm:traversal}
  \textup{\textproc{CheckAll}}$(wl)$ (Algorithm~\ref{alg:subset-space})
  terminates, and if it returns \textup{\textsc{Secure}} then every probe in
  $\llbracket wl \rrbracket$ is secret-independent. In particular
  \textup{\textproc{CheckAll}}$\big\langle (d, \mathcal{W}) \big\rangle$, returning
  \textup{\textsc{Secure}}, certifies $d$-probing security.
  \end{theorem}

  \begin{proof}
  \emph{Termination.} Order worklists by the multiset $\{ |W_1|, \dots, |W_m| \}$
  of their ground-set sizes, under the multiset order. Take any cell the algorithm
  recurses on: it is neither the all-safe cell nor empty, so some factor
  $j^\ast$ has $i_{j^\ast} < c_{j^\ast}$, forcing $B_{j^\ast} \neq \emptyset$;
  and $A_{j^\ast} \supseteq \mathbf{w} \cap W_{j^\ast} \neq \emptyset$ because the
  representative meets every factor. So $W_{j^\ast}$ is replaced by the two
  strictly smaller ground sets $A_{j^\ast}, B_{j^\ast}$, while no other factor's
  ground set grows ($A_j, B_j \subseteq W_j$). The multiset strictly decreases, so
  the recursion is well-founded.

  \emph{Soundness}, by induction along that order. \textproc{CheckAll} returns
  \textsc{Secure} only in one of three ways. If the union $U = \bigcup_j W_j$
  discharges, every subset of $U$ --- hence all of $\llbracket wl \rrbracket$ ---
  is secret-independent by marginalisation. If a degenerate worklist bottoms out,
  $\llbracket wl \rrbracket$ is empty or the lone empty probe, trivially secure.
  Otherwise the representative certifies and the recursive cells all return
  \textsc{Secure}: by Lemma~\ref{lem:set-vandermonde} those cells together with the
  skipped all-safe cell partition $\llbracket wl \rrbracket$; the all-safe cell is
  secret-independent by Component III, and each recursive cell is
  $\llbracket wl_{(i_j)} \rrbracket$ of strictly smaller measure, secret-independent
  by the induction hypothesis. Every probe of $\llbracket wl \rrbracket$ lies in
  exactly one part, so all are secret-independent. Finally
  $\llbracket \langle (d, \mathcal{W}) \rangle \rrbracket = \binom{\mathcal{W}}{d}$,
  and by the size reduction of the section opening a \textsc{Secure} verdict there
  certifies $d$-probing security.
  \end{proof}

  We close this section with two remarks. The soundness above is one-directional by design:
  a \textsc{Secure} verdict is trustworthy, but the traversal inherits Component
  I's incompleteness verbatim and nothing more. An \textsc{Insecure} verdict is
  raised only when Component I declines a representative
  (Section~\ref{sec:general-substitution}), never because a subset slipped through
  unexamined --- Lemma~\ref{lem:set-vandermonde} guarantees the cells cover
  $\llbracket wl \rrbracket$ exactly, so the sweep is genuinely exhaustive. The
  reported witness is therefore a true counterexample precisely when Component I is
  complete on it, as in the linear/affine case; a fuel-bounded refusal there is the
  one place a spurious \textsc{Insecure} can enter, and it enters through Component
  I, not through this traversal. Second, the efficiency of the whole scheme rests
  on the safe set being large: the bigger $\mathsf{safe}$, the fewer and smaller
  the surviving cells, and the closer the union prune comes to settling a branch
  outright. That is exactly the quantity Component III's coupling extension is built
  to maximise, and it is where the approach of this thesis departs from a plain
  set-splitting search.

\section{Component III: Extending Probe Sets}
\label{sec:extension}
\todo[inline]{Introduce the idea of extending a certified probe to certify additional wires cheaply, via closures, replay, and couplings.}
  \subsection{Closure-Driven Extension}
  \label{sec:closures}
  \todo[inline]{Describe extension via closure: wires already implied secret-independent by set-theoretic closure of certified probes.}
  \subsection{Replay-Driven Extension}
  \label{sec:replay}
  \todo[inline]{Describe extension by replaying a certified probe's rewrite derivation against a new wire.}
  \subsection{Coupling-Driven Extension}
  \label{sec:coupling-extension}
  \todo[inline]{Describe extension via the coupling/blinding-certificate view of Section~\ref{sec:coupling-view}: admitting a wire $u$ when $u \circ \varphi_{\mathbf{a}}$ is secret-free.}
