\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Preliminaries}
\label{chap:preliminaries}
% ------------------------------------------------------------
\section{Notation and Mathematical Background}
\label{sec:notation}
filler
  \oldsubsection{\texorpdfstring%
    {The Field $\mathbb{F}_2$ and Multilinear Boolean Polynomials}%
    {The Field F2 and Multilinear Boolean Polynomials}}
  \label{sec:field-f2}
filler
  \subsection{Probability Distributions and Statistical Independence}
  \label{sec:prob-independence}
filler
  \subsection{Boolean Circuits and Gates}
  \label{sec:bool-circuits}

We start, as one should, with the object the rest of the thesis is built on.
The definition is the standard digraph formulation of a circuit, specialised to
arithmetic over the two-element field.

\begin{definition}[Arithmetic circuit over $\mathbb{F}_2$]
\label{def:circuit}
Fix the field $\mathbb{F}_2 = \{0,1\}$ and a set $S = \{s_i\}$ of \emph{gate
functions}, each of some arity $a_i \in \mathbb{N}_{\geq 1}$, i.e.\ each $s_i$
maps $\mathbb{F}_2^{\,a_i}$ to $\mathbb{F}_2$. A \emph{circuit} $C$ is a finite
directed graph (a \emph{digraph}) containing no directed cycle, such that:
\begin{itemize}
  \item Every vertex of indegree $0$ is labelled either by an
  \emph{indeterminate}, in which case it is an \emph{input gate}, or by an
  element of $\mathbb{F}_2$, in which case it is a \emph{constant gate}. The
  input gates are partitioned into \emph{secret} inputs $\mathcal{S}$,
  \emph{random} inputs $\mathcal{R}$, and \emph{public} inputs $\mathcal{P}$.
  \item Every remaining vertex is a \emph{computation gate}, labelled by a
  function $s_i \in S$ whose arity $a_i$ coincides with the vertex's indegree.
  A computation gate of outdegree $0$ is an \emph{output gate}.
\end{itemize}
We call the vertices \emph{wires}, and for a computation gate we speak
interchangeably of the gate and the wire carrying its output.
\end{definition}

Over $\mathbb{F}_2$ there is no difference between the Boolean and the arithmetic
reading of such a circuit: exclusive-or is addition, logical-and is
multiplication, and the two together already express every function
$\mathbb{F}_2^{\,n} \to \mathbb{F}_2$. So we fix
\[
  S = \{\,\oplus,\ \odot\,\},
\]
addition and multiplication in $\mathbb{F}_2$ (Section~\ref{sec:field-f2}), both
of arity $2$. Every computation gate then has indegree exactly $2$ and reads
\[
  w_i = w_j \oplus w_k \qquad \text{or} \qquad w_i = w_j \odot w_k .
\]
Throughout the thesis $\odot$ denotes multiplication in $\mathbb{F}_2$; we
might sometimes use $\cdot$ for the same operation.

Acyclicity is what makes a circuit computable. It lets us list the wires in a
topological order $w_1, \dots, w_m$ in which the two operands of every gate
appear before the gate itself, so that $j, k < i$. Fixing a value in
$\mathbb{F}_2$ for every input gate --- a constant gate carries its own label
--- then determines the value of every wire by induction along this order. When
the inputs are clear from context we write simply $w_i$ for this value.

\begin{notation}
For a wire $w_i$, $\mathrm{vars}(w_i) \subseteq \mathcal{R}$ is the set of
random inputs its defining sub-circuit depends on: $\mathrm{vars}(w_i) = \{w_i\}$
if $w_i \in \mathcal{R}$, $\mathrm{vars}(w_i) = \emptyset$ if
$w_i \in \mathcal{S} \cup \mathcal{P}$ or $w_i$ is a constant gate, and
$\mathrm{vars}(w_i) = \mathrm{vars}(w_j) \cup \mathrm{vars}(w_k)$ for a gate
with operands $w_j, w_k$.
\end{notation}

Two conventions, fixed here once, are used throughout. First, a topological
order realises the circuit as a \emph{straight-line program}: each wire is
defined at most once, in terms only of wires already defined, so there is no
branching and no recomputation of a gate under a different name. We pass freely
between the digraph of Definition~\ref{def:circuit} and this sequential reading,
whichever is clearer at the moment. Second, the restriction to $2$-ary gates is
a modelling choice, not a loss of generality: an $n$-ary sum or product is
always expressible as a chain of $2$-ary gates. Forcing this decomposition up
front matters for Section~\ref{sec:d-probing}, where every gate output is a wire
an adversary may separately observe --- collapsing several operations into one
hidden, unobservable step would understate what a probing adversary can see.
Chapter~\ref{chap:design} (Section~\ref{sec:hash-consed-dag}) later represents
this same object as a hash-consed data structure suited to computation; nothing
here depends on that choice of representation.
\section{Secret Sharing and Masking}
\label{sec:secret-sharing}
filler
  \subsection{Boolean Masking}
  \label{sec:boolean-masking}
filler
\section{The ISW Probing Model}
\label{sec:isw-model}
filler
  \subsection{\NoCaseChange{$t$}-Threshold Probing and Observation Sets}
  \label{sec:t-threshold}
filler
  \subsection{\NoCaseChange{$d$}-Probing Security}
  \label{sec:d-probing}
filler
  \subsection{Probabilistic Non-Interference}
  \label{sec:pni}
filler
  \subsection{Threshold and Strong Non-Interference}
  \label{sec:sni}
filler
\section{Randomness Substitution}
\label{sec:randomness-substitution}

This section isolates, as a single reusable fact, the probabilistic argument
our checker automates. Fix a circuit as in Section~\ref{sec:bool-circuits},
with random inputs $\mathcal{R}$, and let a \emph{randomness assignment} be a
function $\rho : \mathcal{R} \to \{0,1\}$. The secrets and public inputs are
held fixed throughout, so every wire $w$ denotes a function $w(\rho)$ of $\rho$
alone. A \emph{probe} is a tuple of wires --- the values an adversary observes
--- and Section~\ref{sec:d-probing} phrases $d$-probing security as the demand
that the joint distribution of a probe of size at most $d$, taken over the
uniformly distributed $\rho$, not move as the secrets vary.

The whole argument rests on one line of $\mathbb{F}_2$ arithmetic: for any fixed
bit $c$, the translation $b \mapsto c \oplus b$ is a \textbf{bijection} of $\{0,1\}$
that preserves the uniform distribution (i.e. it is its own inverse and merely swaps
the two outcomes). Everything below is this fact, applied to a single coordinate
of $\rho$ with every other coordinate frozen. The object that carries it is a
substitution.

\begin{definition}[Randomness substitution]
\label{def:substitution}
Let $r \in \mathcal{R}$ and let $e$ be a wire with $r \notin \mathrm{vars}(e)$.
The \emph{substitution} $\sigma = (r \mapsto e \oplus r)$ acts on a wire $w$ by
replacing every occurrence of the input $r$ with the wire $e \oplus r$; we write
the resulting wire $w\sigma$. The same $\sigma$ acts on randomness assignments,
through the map $\widehat{\sigma} : \{0,1\}^{\mathcal{R}} \to
\{0,1\}^{\mathcal{R}}$ given by
\[
  \widehat{\sigma}(\rho)(r) = e(\rho) \oplus \rho(r), \qquad
  \widehat{\sigma}(\rho)(r') = \rho(r') \quad \text{for } r' \neq r .
\]
\end{definition}

Geometrically, $\widehat{\sigma}$ is a \emph{shear} of the cube
$\{0,1\}^{\mathcal{R}}$: it fixes every coordinate but $r$ and offsets that one
by $e(\rho)$, a quantity that reads only the frozen coordinates. Restricted to a
fibre --- a pair of points differing solely in the $r$-coordinate --- it is
exactly the translation $\rho(r) \mapsto e(\rho) \oplus \rho(r)$ of the one-line
fact above, so what measure it preserves it preserves by translation invariance,
fibre by fibre. We call any such map --- one coordinate translated by a context
depending on the others, everything else left alone --- a \emph{shear}, and use
the word throughout.

Pairing the two actions under one name is the point of the lemma below: the
syntactic rewrite $w \mapsto w\sigma$ and the reparametrisation
$\rho \mapsto \widehat{\sigma}(\rho)$ are one operation seen from two sides, and
$\widehat{\sigma}$ does not disturb the distribution.

\begin{lemma}[Substitution]
\label{lem:substitution}
For a substitution $\sigma = (r \mapsto e \oplus r)$ as in
Definition~\ref{def:substitution}, the map $\widehat{\sigma}$ is a
measure-preserving involution of the uniform distribution on
$\{0,1\}^{\mathcal{R}}$, and for every wire $w$,
\[
  (w\sigma)(\rho) \;=\; w\big(\widehat{\sigma}(\rho)\big)
  \qquad \text{for every randomness assignment } \rho .
\]
\end{lemma}

\begin{proof}
Fix every coordinate of $\rho$ other than $r$. Since $r \notin \mathrm{vars}(e)$,
the value $e(\rho)$ depends only on these frozen coordinates, so it is constant
as $\rho(r)$ ranges over $\{0,1\}$. On this fibre $\widehat{\sigma}$ acts by the
translation $\rho(r) \mapsto e(\rho) \oplus \rho(r)$, which is measure-preserving
and self-inverse by the observation above, while every other coordinate is left
fixed. Hence $\widehat{\sigma}$ is an involution of $\{0,1\}^{\mathcal{R}}$
preserving the uniform measure coordinatewise, and so preserving the product
measure.

The identity is a structural induction on $w$. For $w = r$ both sides equal
$e(\rho) \oplus \rho(r)$. For an input $w \neq r$ both sides equal $w(\rho)$,
since $\widehat{\sigma}$ fixes every coordinate but $r$. For a gate
$w_i = w_j \oplus w_k$ or $w_i = w_j \odot w_k$, apply the hypothesis to $w_j$
and $w_k$ and use that $\oplus$ and $\odot$ are computed pointwise from their
operands' values.
\end{proof}

\begin{corollary}[Single-occurrence discharge]
\label{cor:single-occurrence-discharge}
Suppose $r$ occurs exactly once among a probe $\mathbf{w}$ and the wires it
depends on, as an operand of a gate $x = e \oplus r$ with
$r \notin \mathrm{vars}(e)$, and let $\sigma = (r \mapsto e \oplus r)$. Then
$\mathbf{w}\sigma$ is identically distributed to $\mathbf{w}$ over $\rho$
uniform, and every occurrence of $x$ in $\mathbf{w}\sigma$ has collapsed to the
bare input $r$.
\end{corollary}

\begin{proof}
Distributional equality is Lemma~\ref{lem:substitution} applied coordinatewise
to $\mathbf{w}$, using that $\widehat{\sigma}$ preserves the uniform
distribution on $\{0,1\}^{\mathcal{R}}$. For the second claim,
$x\sigma = e \oplus (e \oplus r) = r$ by $\mathbb{F}_2$-cancellation; since $r$
occurs nowhere else by hypothesis, no other part of $\mathbf{w}$ is touched.
\end{proof}

\begin{remark}
Rewriting a probe by a substitution --- whether in the general form of
Lemma~\ref{lem:substitution} or the single-occurrence special case of
Corollary~\ref{cor:single-occurrence-discharge} --- leaves its distribution
untouched, hence leaves untouched whether it depends on the secrets. That is the
whole lever: to decide security of a probe it is enough to rewrite it, one
substitution at a time, into a form with no secret in sight. The
single-occurrence case is the one the \texttt{maskVerif} tool
\parencite{10.1007/978-3-030-29959-0_15} calls \emph{optimistic sampling}, and
the name is apt --- Corollary~\ref{cor:single-occurrence-discharge} never
inspects the content of $e$, firing whether or not $e$ depends on the secrets,
and checking only the syntactic pattern that $r$ occurs once, additively. A
derivation is nothing more than such substitutions chained together; how that
chain is searched for, and what the composite of its $\widehat{\sigma}$
witnesses, is taken up in the later chapters.
\end{remark}

\section{Computational Hardness Preview}
\label{sec:hardness-preview}
filler
