\setlength{\parindent}{0pt}
% ------------------------------------------------------------
\chapter{Related Work}
% ------------------------------------------------------------
\section{Language-Based Verification} % main inspiration
\todo[inline]{Survey language-based approaches to masking verification, the main inspiration for this thesis.}
  \subsection{Verified Proofs of Higher-Order Masking}
  \todo[inline]{Discuss prior work on verified/formal proofs of higher-order masking schemes.}
  \subsection{Large Sets and the Exploration Algorithm}
  \todo[inline]{Discuss prior large-set exploration algorithms for probing security and how this thesis's subset-space traversal relates.}
  \subsection{Witness Replay and Derivation Reuse}
  \todo[inline]{Discuss prior work on reusing a certification witness/derivation across related probes, and how it relates to Component III's replay-driven extension (Section~\ref{sec:replay}).}
  \subsection{maskVerif with Physical Defaults}
  \todo[inline]{Discuss \texttt{maskVerif} and its physical-default extensions, and compare to the rewrite engine of Chapter~\ref{chap:design}.}
\section{Fourier-Analytic Verification} % Bloem
\label{sec:fourier-analytic}
\todo[inline]{Survey Fourier-analytic approaches to masking verification (Bloem et al.).}
  \subsection{The Xiao-Massey Criterion and Approximations}
  \todo[inline]{Discuss the Xiao-Massey criterion and its approximate variants for testing correlation-immunity/independence.}
  \subsection{SAT-Based Realization and Its Cost}
  \todo[inline]{Discuss SAT-based realizations of Fourier-analytic verification and their computational cost.}
\section{Other Approaches}
\todo[inline]{Survey other verification approaches not covered above.}
  \subsection{Model Counting and SMT}
  \todo[inline]{Discuss model-counting and SMT-based approaches to verifying secret-independence.}
  \subsection{Compositional Reasoning and Relabeling}
  \todo[inline]{Discuss compositional reasoning and relabeling/symmetry techniques used elsewhere, relating to Future Work's symmetry-orbit proposal.}
\section{Position of This Work}
\todo[inline]{Position this thesis's contributions relative to the surveyed related work.}
